---
title: "ProcessGEEZonalStats"
author: "Eric Jensen"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(mapview)
library(reshape2)
library(SparkR)
```

### Import zonal statistics CSVs and process to prepare for join
```{r}
# list the csv files for each zonal statistic
ndvi_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNDVI", full.names = T)
ndvi_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNDVI", full.names = T)
nbr_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNBR", full.names = T)
nbr_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNBR", full.names = T)
tcb_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCB", full.names = T)
tcb_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCB", full.names = T)
tcg_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCG", full.names = T)
tcg_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCG", full.names = T)
tcw_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCW", full.names = T)
tcw_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCW", full.names = T)
listofLists = list(ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

# create empty lists for each statistic
ndvi_mean <- list()
ndvi_std <- list()
nbr_mean <- list()
nbr_std <- list()
tcb_mean <- list()
tcb_std <- list()
tcg_mean <- list()
tcg_std <- list()
tcw_mean <- list()
tcw_std <- list()

# functions to read the CSVs into the list
ndvi_mean <- map(ndvi_mean_zonal,read_csv)
ndvi_std <- map(ndvi_std_zonal,read_csv)
nbr_mean <- map(nbr_mean_zonal,read_csv)
nbr_std <- map(nbr_std_zonal,read_csv)
tcb_mean <- map(tcb_mean_zonal,read_csv)
tcb_std <- map(tcb_std_zonal,read_csv)
tcg_mean <- map(tcg_mean_zonal,read_csv)
tcg_std <- map(tcg_std_zonal,read_csv)
tcw_mean <- map(tcw_mean_zonal,read_csv)
tcw_std <- map(tcw_std_zonal,read_csv)
remove(listofLists,ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

#bind cols to produce dataframes of each zonal statistic, rename cols with unique name
ndvi_mean <- ndvi_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_mean = mean)
ndvi_std <- ndvi_std %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_stdDev = stdDev)
nbr_mean <- nbr_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_mean = mean)
nbr_std <- nbr_std %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_stdDev = stdDev)
tcb_mean <- tcb_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_mean = mean)
tcb_std <- tcb_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_stdDev = stdDev)
tcg_mean <- tcg_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_mean = mean)
tcg_std <- tcg_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_stdDev = stdDev)
tcw_mean <- tcw_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_mean = mean)
tcw_std <- tcw_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_stdDev = stdDev)

# Assessed all dataframes for missing predictor variable values, none present
```

# Import spatial plots and join with GEE zonal stats
```{r}
# Import plots sf for joining GEE zonal stats too
AIM_plots <- read_sf('data/AIM_model.shp') %>%
  st_transform(crs = 4326)
AIMdf <- as_tibble(matrix(unlist(AIM_plots), nrow=nrow(AIM_plots), byrow=F))
AIM_plotsDF <- dplyr::rename(AIMdf, "plotkey" = V1, "PlotID" = V2, "VisitDate" = V3, "VisitYear" = V4, "PrimaryKey" = V5, "SpeciesLis" = V6, "geometry" = V7)
View(AIMdf)
View(AIM_plots)

listofGEE = list(AIM_plots,ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std)
# Join the zonal statistics to each plot, remove plots with NAs--likely only the dateless plot
AIM_wGEE <- reduce(listofGEE, left_join)
remove(ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std, listofGEE)
```

### Calculate zonal statistics for rasters of elevation, slope, aspect, etc.
```{r}
# import physical rasters
GBdem<- raster('data/Tiffs/GBdem_new.tif')
GBslope<- raster('data/Tiffs/GBslope.tif')
GBaspect<- raster('data/Tiffs/GBaspect.tif')
GBtpi<- raster('data/Tiffs/GBtpi.tif')

# extract raster values and mutate new column to prediction layers
AIM_wGEE_wPhys <- AIM_wGEE %>%
  mutate(elev = extract(x=GBdem, y=AIM_wGEE)) %>%
  mutate(aspect = extract(x=GBaspect, y=AIM_wGEE)) %>%
  mutate(slope = extract(x=GBslope, y=AIM_wGEE)) %>%
  mutate(tpi = extract(x=GBtpi, y=AIM_wGEE))
remove(GBdem,GBslope,GBaspect,GBtpi)
```

### Calculate species richness values based on list
```{r}
# Parse the species list strings to lists
AIM_wGEE_wPhys_wSR <- AIM_wGEE_wPhys %>%
  mutate(SpecList = str_split(SpeciesLis,pattern=";")) %>%
  dplyr::select(-SpeciesLis)

# Calculate number of species at each plot (-1 because of extra "") and bind the dataframe to predictive dataframe
SpecN = list()
for (i in 1:nrow(AIM_wGEE_wPhys)) {
          SpecN[i] <- length(AIM_wGEE_wPhys_wSR$SpecList[[i]]) - 1}
SpecNdf <- data.frame(matrix(unlist(SpecN), nrow=length(SpecN), byrow=T))
AIM_wGEE_wPhys_wSR <- bind_cols(AIM_wGEE_wPhys_wSR,SpecNdf)

AIM_PhysRS <- AIM_wGEE_wPhys_wSR
remove(AIM_wGEE,AIM_wGEE_wPhys,AIM_wGEE_wPhys_wSR)

# Write existing table to CSV
AIM_df <- data.frame(matrix(unlist(AIM_PhysRS), nrow=length(AIM_PhysRS), byrow=T))
write_csv(AIMcovariates, path = "data/CSVs/AIMcovariates.csv")

AIMcovariates$VisitYear
# Read in USDA plants dataframe to get genus and species data
# plantsList <- read_csv('data/TXT/USDA_PlantsTable.txt')
```

### Assess climate covariate data (36 deg - 45 deg; 112 deg - 122 deg)
#### Climate variables of interest are:
- ADI (annual dryness index)
- SDI (summer dryness index)
- FFP (frost free period)
- GSP (growing season precipitation)
- MAP (mean annual precipitation)
- SPRP (spring precipitation)
- SMRP (summer precipitation)
```{r}
# import CSV from previous lines

# Import mosaics of desired variables
cliIndexList <- c("_adi.tif", "_sdi.tif", "_sprp.tif", "_smrp.tif", "_ffp.tif", "_gsp.tif", "_map.tif")
cliFolderList <- list()
cliFolderList <- grep(list.files(path = 'D:/current_climates', full.names = T), pattern = "33|34|35|111|123|124", invert = T, value = T)

# only create the list once--otherwise it will overwrite
cliList <- list() 
cliMosaic <- list()

# vary the index from 1:the number of indices
index = 2
cliIndexList[[index]]
cliList[[index]] <- list()
# cliList[[index]]$index <- str_replace(cliIndexList[[index]], ".tif", "")
# cliList[[index]]$index <- str_replace(cliList[[index]]$index, "_", "")

for(i in seq_along(cliFolderList)){ 
    cliList[[index]][[i]] <- raster(grep(list.files(path = cliFolderList[[i]], full.names = T), 
                            pattern = cliIndexList[[index]], value = T))}
cliList[[index]]$fun <- mean

cliMosaic[[index]] <- do.call(mosaic,cliList[[index]])
names(cliMosaic[[index]]) <- cliIndexList[[index]] %>% str_replace(".tif", "") %>% str_replace("_", "")

cliMosaic


# cliMosaicList <- list()
# cliMosaicList$fun <- mean
# rasterList <- list()
# rasterList$fun <- mean
# 
# for(i in seq_along(cliIndexList)){
#   for(j in 1:2){ #seq_along(cliFolderList)
#     rasterList[[j]] <- raster(grep(list.files(path = cliFolderList[[j]],full.names = T), 
#                             pattern = cliIndexList[[i]], value = T))
#     rasterList$fun <- mean
#     cliMosaicList[[j]] <- do.call(mosaic, rasterList)
#     cliMosaicList$fun <- mean
#     }
#   }
```

### Assess soils data
```{r}

```

### Review biophysical data
```{r}

```







