---
title: "ProcessGEEZonalStats"
author: "Eric Jensen"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(mapview)
library(raster)
library(spatialEco)
library(randomForest)
library(rfUtilities)
library(corrplot)
library(ggthemes)
library(egg)
library(ggpubr)
library(spatstat)
library(tmaptools)
```

### Import zonal statistics CSVs and process to prepare for join
```{r}
# list the csv files for each zonal statistic
ndvi_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNDVI", full.names = T)
ndvi_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNDVI", full.names = T)
nbr_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNBR", full.names = T)
nbr_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNBR", full.names = T)
tcb_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCB", full.names = T)
tcb_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCB", full.names = T)
tcg_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCG", full.names = T)
tcg_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCG", full.names = T)
tcw_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCW", full.names = T)
tcw_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCW", full.names = T)
listofLists = list(ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

RScovariates <- length(listofLists)

# create empty lists for each statistic
ndvi_mean <- list()
ndvi_std <- list()
nbr_mean <- list()
nbr_std <- list()
tcb_mean <- list()
tcb_std <- list()
tcg_mean <- list()
tcg_std <- list()
tcw_mean <- list()
tcw_std <- list()

# functions to read the CSVs into the list
ndvi_mean <- map(ndvi_mean_zonal,read_csv)
ndvi_std <- map(ndvi_std_zonal,read_csv)
nbr_mean <- map(nbr_mean_zonal,read_csv)
nbr_std <- map(nbr_std_zonal,read_csv)
tcb_mean <- map(tcb_mean_zonal,read_csv)
tcb_std <- map(tcb_std_zonal,read_csv)
tcg_mean <- map(tcg_mean_zonal,read_csv)
tcg_std <- map(tcg_std_zonal,read_csv)
tcw_mean <- map(tcw_mean_zonal,read_csv)
tcw_std <- map(tcw_std_zonal,read_csv)
remove(listofLists,ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

#bind cols to produce dataframes of each zonal statistic, rename cols with unique name
ndvi_mean <- ndvi_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_mean = mean)
ndvi_std <- ndvi_std %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_stdDev = stdDev)
nbr_mean <- nbr_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_mean = mean)
nbr_std <- nbr_std %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_stdDev = stdDev)
tcb_mean <- tcb_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_mean = mean)
tcb_std <- tcb_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_stdDev = stdDev)
tcg_mean <- tcg_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_mean = mean)
tcg_std <- tcg_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_stdDev = stdDev)
tcw_mean <- tcw_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_mean = mean)
tcw_std <- tcw_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_stdDev = stdDev)

# Assessed all dataframes for missing predictor variable values, none present
```

# Import spatial plots and join with GEE zonal stats
```{r}
# Import plots sf for joining GEE zonal stats too
AIM_plots <- read_sf('data/AIM_model.shp') %>%
  st_transform(crs = 4326)
# Convert to tibble without geometry for manipulating and joining
AIMdf <- as_data_frame(matrix(unlist(AIM_plots), nrow=nrow(AIM_plots), byrow=F))
AIM_plotsDF <- AIMdf %>%
  dplyr::select(-V3, -V7, -V8) %>%
  dplyr::rename( "plotkey" = V1, "PlotID" = V2, "VisitYear" = V4, 
                 "PrimaryKey" = V5, "SpeciesLis" = V6)

listofGEE = list(AIM_plotsDF,ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std)
# Join the zonal statistics to each plot, remove plots with NAs--likely only the dateless plot
AIM_wGEE <- reduce(listofGEE, left_join)
remove(ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std, listofGEE, AIMdf)
```

### Calculate zonal statistics for rasters of elevation, slope, aspect, etc.
```{r}
# AIM_buffers <- read_sf('data/AIM_modelbuff_100m.shp') %>%
#   st_transform(crs = 4326)

# bind primary key to join by
# AIM_wGEE <- AIM_wGEE %>%
#   left_join(AIM_plots, "PrimaryKey") %>%
#   dplyr::select(-ends_with(".y"))
# 
# # subset remote sensing covariate columns to bind to sf object
# to_join<- as_tibble(AIM_wGEE[6:(5+RScovariates)])
# AIM_RS = bind_cols(AIM_plots,to_join)
# View(AIM_RS)

### Produce and export biophysical rasters
#GBslope <- terrain(GBdem, opt = 'slope', unit = 'degrees')
# writeRaster(GBslope, filename = 'data/Tiffs/GBslope.tif', format = "GTiff")
#GBaspect <- terrain(GBdem, opt = 'aspect', unit = 'degrees')
#writeRaster(GBaspect, filename = 'data/Tiffs/GBaspect.tif', format = "GTiff")
#GBtpi <- terrain(GBdem, opt = 'tpi', unit = 'degrees')
#writeRaster(GBtpi, filename = 'data/Tiffs/GBtpi.tif', format = "GTiff")
#GBtrasp <- trasp(GBdem)
# GBhli <- hli(GBdem)
# writeRaster(GBtrasp, filename = 'data/Tiffs/GBtrasp.tif', format = "GTiff")
# writeRaster(GBhli, filename = 'data/Tiffs/GBhli.tif', format = "GTiff")

### Import physical rasters
# GBdem<- raster('data/Tiffs/GBdem_new.tif')
# GBslope<- raster('data/Tiffs/GBslope.tif')
# GBaspect<- raster('data/Tiffs/GBaspect.tif')
# GBtpi<- raster('data/Tiffs/GBtpi.tif')
# GBtrasp<- raster('data/Tiffs/GBtrasp.tif')
# GBhli<- raster('data/Tiffs/GBhli.tif')
#GBlinaspect<- raster('data/Tiffs/GBlinaspect.tif')

### extract raster values and mutate new column to prediction layers
# AIM_buffers_Phys <- as_tibble(AIM_buffers) %>%
#   dplyr::mutate(elev = raster::extract(x=GBdem, y=AIM_buffers, fun=mean)) %>%
#   dplyr::mutate(aspect = raster::extract(x=GBaspect, y=AIM_buffers, fun=mean)) %>%
#   dplyr::mutate(slope = raster::extract(x=GBslope, y=AIM_buffers, fun=mean)) %>%
#   dplyr::mutate(tpi = raster::extract(x=GBtpi, y=AIM_buffers, fun=mean))

# AIM_covariates_cli_soil_physwrk <- as_tibble(AIM_covariates_cli_soil) %>%
#   dplyr::mutate(hli = raster::extract(x=GBhli, y=AIM_covariates_cli_soil, fun=mean)) %>%
#   dplyr::mutate(linaspect = raster::extract(x=GBlinaspect, y=AIM_covariates_cli_soil, fun=mean)) %>%
#   dplyr::mutate(trasp = raster::extract(x=GBtrasp, y=AIM_covariates_cli_soil, fun=mean))

# # Get latitude as covariate
# AIM_plots <- read_sf('data/PlotsShapefiles/AIM_model.shp') %>%
#   st_transform(crs = 4326)
# AIM_plots <- AIM_plots %>%
#     mutate(long = unlist(map(AIM_plots$geometry,1)),
#            lat = unlist(map(AIM_plots$geometry,2)))
# AIM_covariates_wlat <- bind_cols(AIM_covariates,select(AIM_plots, lat))
# View(AIM_covariates_wlat)
# st_write(AIM_covariates_wlat, 'data/Working/AIM_covariates_working_final.shp')
```

### Calculate species richness values based on list
```{r}
# Parse the species list strings to lists
AIM_RS_Phys_SR <- AIM_RS_Phys %>%
  dplyr::mutate(SpecList = str_split(SpeciesLis,pattern=";")) %>%
  dplyr::select(-SpeciesLis)
View(AIM_RS_Phys_SR)

# Calculate number of species at each plot (-1 because of extra "") and bind the dataframe to predictive dataframe
SpecN = list()
for (i in 1:nrow(AIM_RS_Phys_SR)) {
          SpecN[i] <- length(AIM_RS_Phys_SR$SpecList[[i]]) - 1}
SpecNdf <- data.frame(matrix(unlist(SpecN), nrow=length(SpecN), byrow=T))
AIM_RS_Phys_SR_count <- bind_cols(AIM_RS_Phys_SR,SpecNdf)

View(AIM_RS_Phys_SR_count)
AIM_PhysRS <- AIM_wGEE_wPhys_wSR
remove(AIM_wGEE,AIM_wGEE_wPhys,AIM_wGEE_wPhys_wSR)

# Write existing table to CSV
st_write(AIM_covariates, 'data/Working/AIM_covariates.shp')

# Read in USDA plants dataframe to get genus and species data
# plantsList <- read_csv('data/TXT/USDA_PlantsTable.txt')
```


### Read-in output CSV above to avoid running everything upstream
```{r}
AIM_working <- read_sf('data/Working/AIM_covariates.shp') %>%
  st_transform(crs = 4326)

# View(AIM_working)
# mapview(AIM_working)
```


### Assess climate covariate data (36 deg - 45 deg; 112 deg - 122 deg)
#### Climate variables of interest are:
- ADI (annual dryness index)
- SDI (summer dryness index)
- FFP (frost free period)
- GSP (growing season precipitation)
- MAP (mean annual precipitation)
- SPRP (spring precipitation)
- SMRP (summer precipitation)
```{r}
# ### Import CSV from previous lines
# 
# ### Import mosaics of desired variables
# cliIndexList <- c("_adi.tif", "_sdi.tif", "_sprp.tif", "_smrp.tif", "_ffp.tif", "_gsp.tif", "_map.tif")
# cliFolderList <- list()
# cliFolderList <- grep(list.files(path = 'D:/current_climates', full.names = T), pattern = "33|34|35|111|123|124", invert = T, value = T)
# 
# ### Only create the list once--otherwise it will overwrite
# cliList <- list() 
# cliMosaic <- list()
# 
# ### Vary the index from 1:the number of indices
# index = 2
# cliIndexList[[index]]
# cliList[[index]] <- list()
# # cliList[[index]]$index <- str_replace(cliIndexList[[index]], ".tif", "")
# # cliList[[index]]$index <- str_replace(cliList[[index]]$index, "_", "")
# 
# ### Add each image into a list of images for a given index
# for(i in seq_along(cliFolderList)){ 
#     cliList[[index]][[i]] <- raster(grep(list.files(path = cliFolderList[[i]], full.names = T), 
#                             pattern = cliIndexList[[index]], value = T))}
# #View(cliList)
# cliList[[index]]$fun <- mean
# 
# 
# #-------------------------------------------------------------------------
# ### Some indices are missing some images--this code runs for those indices
# parse1 = paste(cliFolderList, '/', sep = "")
# parse2 = paste(parse1,(substr(cliFolderList,21,31)), sep = "")
# sdilist <- paste(parse2, cliIndexList[index], sep = "")
# print(sdilist)
# 
# sdiraster <- vector(mode = "list", length = 99)
# 
# # Create list of Raster files (should be cleaning it as well but doesn't seem to want to)
# for(i in seq_along(sdilist)){
#   if(file.exists(sdilist[i]) == T)
#     sdiraster[[i]] <- raster(sdilist[i])
# }
# # Clean the list by removing NULL values
# for(i in 1:length(sdiraster)){
#   if(is.null(sdiraster[[i]])==TRUE)
#     sdiraster[[i]] <- NULL
# }
# 
# for(i in 1:length(sdiraster)){
#   print(is.null(sdiraster))}
# 
# # Necessary to apply the do.call
# names(sdiraster) <- NULL
# sdiraster$fun <- mean
# # Run the do.call over the list of items to mosaic
# sdiMosaic <- do.call(mosaic,sdiraster)
# 
# names(cliMosaic[[index]]) <- cliIndexList[[index]] %>% str_replace(".tif", "") %>% str_replace("_", "")
# 
# #-------------------------------------------------------------------------
# ### Mosaic images together for each index
# cliList[[index]]$fun <- mean
# 
# cliMosaic[[index]] <- do.call(mosaic,cliList[[index]])
# names(cliMosaic[[index]]) <- cliIndexList[[index]] %>% str_replace(".tif", "") %>% str_replace("_", "")
# 
# View(cliMosaic[[index]])
# plot(cliMosaic[[index]])
# 
# #-------------------------------------------------------------------------
# ### Write rasters
# 
# writeRaster(adiMosaic, filename = 'data/Tiffs/climate_adi.tif', format = "GTiff")
# writeRaster(sdiMosaic, filename = 'data/Tiffs/climate_sdi.tif', format = "GTiff")
# writeRaster(cliMosaic[[3]], filename = 'data/Tiffs/climate_sprp.tif', format = "GTiff")
# writeRaster(cliMosaic[[4]], filename = 'data/Tiffs/climate_smrp.tif', format = "GTiff")
# writeRaster(cliMosaic[[5]], filename = 'data/Tiffs/climate_ffp.tif', format = "GTiff")
# writeRaster(cliMosaic[[6]], filename = 'data/Tiffs/climate_gsp.tif', format = "GTiff")
# writeRaster(cliMosaic[[7]], filename = 'data/Tiffs/climate_map.tif', format = "GTiff")

```

### Calculate zonal statistics for each climate variable of interest
```{r}
# ### Import climate mosaics
# Climate_adi <- raster('data/Tiffs/climate_adi.tif')
# Climate_sdi <- raster('data/Tiffs/climate_sdi.tif')
# Climate_sprp <- raster('data/Tiffs/climate_sprp.tif')
# Climate_smrp <- raster('data/Tiffs/climate_smrp.tif')
# Climate_ffp <- raster('data/Tiffs/climate_ffp.tif')
# Climate_gsp <- raster('data/Tiffs/climate_gsp.tif')
# Climate_map <- raster('data/Tiffs/climate_map.tif')
# 
# AIM_working_cli <- as_tibble(AIM_working) %>%
#   dplyr::mutate(Cli_adi = raster::extract(x=Climate_adi, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_sdi = raster::extract(x=Climate_sdi, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_sprp= raster::extract(x=Climate_sprp, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_smrp= raster::extract(x=Climate_smrp, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_ffp = raster::extract(x=Climate_ffp, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_gsp = raster::extract(x=Climate_gsp, y=AIM_working, fun=mean)) %>%
#   dplyr::mutate(Cli_map = raster::extract(x=Climate_map, y=AIM_working, fun=mean))
# 
# # write the new table to shapefile
# st_write(AIM_working_cli, 'data/Working/AIM_covariates_cli.shp')
```

### Mosaic soils data and export mosaics
```{r}
# library(raster)
# SoilWater <- list()
# SoilWater$fun <- mean
# SoilWater[[1]] <- raster('data/Tiffs/Soil/SoilWater1.tif')
# SoilWater[[2]] <- raster('data/Tiffs/Soil/SoilWater2.tif')
# SoilWater[[3]] <- raster('data/Tiffs/Soil/SoilWater3.tif')
# SoilWater[[4]] <- raster('data/Tiffs/Soil/SoilWater4.tif')
# names(SoilWater) <- NULL
# SoilWater$fun <- mean
# SoilWaterMosaic <- do.call(mosaic,SoilWater)
# 
# SoilCarbon <- list()
# SoilCarbon[[1]] <- raster('data/Tiffs/Soil/Carbon1.tif')
# SoilCarbon[[2]] <- raster('data/Tiffs/Soil/Carbon2.tif')
# SoilCarbon[[3]] <- raster('data/Tiffs/Soil/Carbon3.tif')
# SoilCarbon[[4]] <- raster('data/Tiffs/Soil/Carbon4.tif')
# names(SoilCarbon) <- NULL
# SoilCarbon$fun <- mean
# SoilCarbonMosaic <- do.call(mosaic,SoilCarbon)
# 
# SoilTexture <- list()
# SoilTexture[[1]] <- raster('data/Tiffs/Soil/TextureClass1.tif')
# SoilTexture[[2]] <- raster('data/Tiffs/Soil/TextureClass2.tif')
# SoilTexture[[3]] <- raster('data/Tiffs/Soil/TextureClass3.tif')
# SoilTexture[[4]] <- raster('data/Tiffs/Soil/TextureClass4.tif')
# names(SoilTexture) <- NULL
# SoilTexture$fun <- mean
# SoilTextureMosaic <- do.call(mosaic,SoilTexture)
# 
# SoilSand <- list()
# SoilSand[[1]] <- raster('data/Tiffs/Soil/Clay1.tif')
# SoilSand[[2]] <- raster('data/Tiffs/Soil/Clay2.tif')
# SoilSand[[3]] <- raster('data/Tiffs/Soil/Clay3.tif')
# names(SoilSand) <- NULL
# SoilSand$fun <- mean
# SoilSandMosaic <- do.call(mosaic,SoilSand)
# 
# SoilClay <- list()
# SoilClay[[1]] <- raster('data/Tiffs/Soil/Sand1.tif')
# SoilClay[[2]] <- raster('data/Tiffs/Soil/Sand2.tif')
# SoilClay[[3]] <- raster('data/Tiffs/Soil/Sand3.tif')
# SoilClay[[4]] <- raster('data/Tiffs/Soil/Sand4.tif')
# names(SoilClay) <- NULL
# SoilClay$fun <- mean
# SoilClayMosaic <- do.call(mosaic,SoilClay)
# 
# writeRaster(SoilWaterMosaic,  filename = 'data/Tiffs/Soil/SoilWaterMosaic.tif', format = "GTiff")
# writeRaster(SoilCarbonMosaic, filename = 'data/Tiffs/Soil/SoilCarbonMosaic.tif', format = "GTiff")
# writeRaster(SoilTextureMosaic,filename = 'data/Tiffs/Soil/SoilTextureMosaic.tif', format = "GTiff")
# writeRaster(SoilSandMosaic,   filename = 'data/Tiffs/Soil/SoilSandMosaic.tif', format = "GTiff")
# writeRaster(SoilClayMosaic,   filename = 'data/Tiffs/Soil/SoilClayMosaic.tif', format = "GTiff")
```

### Calculate zonal statistics for each soils variable of interest
```{r}
### Read-insoils mosaics
SoilWaterMosaic <- raster('data/Tiffs/Soil/SoilWaterMosaic.tif')
SoilCarbonMosaic <- raster('data/Tiffs/Soil/SoilCarbonMosaic.tif')
SoilTextureMosaic <- raster('data/Tiffs/Soil/SoilTextureMosaic.tif')
SoilSandMosaic <- raster('data/Tiffs/Soil/SoilSandMosaic.tif')
SoilClayMosaic <- raster('data/Tiffs/Soil/SoilClayMosaic.tif')

AIM_working_cli_soil <- as_tibble(AIM_working) %>%
  dplyr::mutate(soil_h2o = raster::extract(x=SoilWaterMosaic, y=AIM_working, fun=mean)) %>%
  dplyr::mutate(soil_C = raster::extract(x=SoilCarbonMosaic, y=AIM_working, fun=mean)) %>%
  dplyr::mutate(soil_tex = raster::extract(x=SoilTextureMosaic, y=AIM_plots, fun=mean)) %>% #needs to be points, because classes
  dplyr::mutate(soil_snd = raster::extract(x=SoilSandMosaic, y=AIM_working, fun=mean)) %>%
  dplyr::mutate(soil_cly = raster::extract(x=SoilClayMosaic, y=AIM_working, fun=mean))

# Write to shapefile once zonal statistics have been calculated
st_write(AIM_working_cli_soil, 'data/Working/AIM_covariates_cli_soil.shp')
```

########### MASK FORESTS BEFORE PREDICTIVE MASKS ###############
########## Move modeling portion to another script #############

################## Random Forest Modeling ######################
### Model Selection

### Clean covariate data for modeling
```{r}
### Assess missing values and anything else to clean up
AIM_covariates <- read_sf('data/Working/AIM_covariates_model.shp') %>%
  st_transform(crs = 4326)
AIM_covariates <- st_centroid(AIM_covariates)
# Remove columns not needed for modeling
AIM_covariates <- dplyr::select(AIM_covariates, -c('PlotKey','VisitDate', 'PrimaryKey', 'SpeciesLis', 'sl_tx_c')) #removed soil texture class because it's a factor, not numeric

# Remove plots with NAs for one or more variable 
AIM_model <- AIM_covariates %>%
  drop_na()
```

```{r}
### Generate random number to set seed
# sample(1:1000000,1)
rf_seed = 517953
set.seed(rf_seed)

# Put in base R dataframe for RF model
AIM_rf <- data.frame(AIM_model[,3],AIM_model[,4:31])
#View(AIM_rf)

# ##### Subset data to test model selection and model fit #####
# # Model selection
# AIM_rf_subs <- slice(AIM_rf,1:500)
# rfselSubs <- rf.modelSel(xdata = AIM_rf_subs[,2:29], ydata = AIM_rf_subs[,1], r = c(0.1, 0.2, 0.5, 0.7, 0.9))
# rfselSubs$selvars
# 
# # Model fitting
# rfselDF_sub <- data.frame(AIM_rf_subs[,1], AIM_rf_subs[c("elev", "ndvi_mn", "ndv_stD", "nbr_men", "tcg_men", "tcg_stD", "Cli_adi", "Cli_sdi", "Cl_sprp", "Cl_smrp", "Cli_ffp", "Cli_gsp", "Cli_map", "lat")])
# 
# rf_fit_sub = randomForest(x = rfselDF_sub[2:15], y= rfselDF_sub[,1])
# print(rf_fit_sub)

##### Run on full dataset #####
View(AIM_rf)
predictors = AIM_rf[,2:29]
response = AIM_rf[,1]
rfselFull = rf.modelSel(xdata = predictors, ydata = response, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) # Need to assess parsimony in future models

plot(rfselFull)

rfselFull <- cbind(rownames(rfselFull$sel.importance), rfselFull$sel.importance)
rownames(rfselFull) <- NULL
colnames(rfselFull) <- c("name", "imp")

rfselFull_ord <- rfselFull[order(-rfselFull$imp),]
nrow(rfselFull_ord)
# drop columns that are not as important based on rfmodelsel

raster_cov_names <- rfselFull_ord$name[1:14]

rastercolumns_cov <- predictors[,as.character(raster_cov_names)]#[-c(4,5,6,9,8,7)]]

# calculate correlation coefficient matrix
correlation <-cor(rastercolumns_cov, method="pearson")

# plot the correlation. the darker the number, the more correlated the two
# variables
corrplot(correlation,method="number")

rastercolumns_full <- predictors[,as.character(raster_cov_names)]
rastercolumns_6var <- predictors[,as.character(raster_cov_names)[-c(7:14)]]
rastercolumns_7var <- predictors[,as.character(raster_cov_names)[-c(7,8,9,11,12,13,14)]]

# calculate correlation coefficient matrix for the variables for the two models
correlation6 <-cor(rastercolumns_6var, method="pearson")
correlation7 <-cor(rastercolumns_7var, method="pearson")

# plot the correlation. the darker the number, the more correlated the twovariables
corrplot(correlation6,method="number")
corrplot(correlation7,method="number")

rfselFull$selvars
## You can now manually remove any variables that are above your correlation 
## threshold. In this case we will just leave them all in, but we could remove
## them by using the following code

###### Model fitting
data_cov_model6 <- cbind(rastercolumns_6var, specN= response)
data_cov_model7 <- cbind(rastercolumns_7var, specN= response)
data_cov_modelFull <- cbind(rastercolumns_full, specN= response)

# Now let's restrict our final predictors to those that are most important (Change manually) 
# and run the continuous model

rf_model6 = randomForest(specN ~ ., data=data_cov_model6, importance = TRUE, ntree = 3000, mtry = 2)
rf_model7 = randomForest(specN ~ ., data=data_cov_model7, importance = TRUE, ntree = 3000, mtry = 2)
rf_modelFull = randomForest(specN ~ ., data=data_cov_modelFull, importance = TRUE, ntree = 3000, mtry = 2)

model6varEx <- round(rf_model6$rsq[3000], digits = 4)
model7varEx <- round(rf_model7$rsq[3000], digits = 4)
modelFullvarEx <- round(rf_modelFull$rsq[3000], digits = 4)

# 
# # Model fitting
# rfselDF <- data.frame(AIM_rf[,1], AIM_rf[c("elev", "slope", "ndvi_mn", "tcb_men", "tcg_men", "Cli_adi", "Cli_sdi", "Cl_sprp", "Cli_ffp", "Cli_map", "soil_h2", "sol_cly", "trasp", "lat")])
# varsel_cor <- cor(AIM_rf[c("elev", "slope", "ndvi_mn", "tcb_men", "tcg_men", "Cli_adi", "Cli_sdi", "Cl_sprp", "Cli_ffp", "Cli_map", "soil_h2", "sol_cly", "trasp", "lat")])
# corrplot(varsel_cor, method = "number")
# 
# rf_fit <- list()
# rf_fit[[3]] = randomForest(x = rfselDF[2:length(rfselDF)], y= rfselDF[,1], ntree = 3000, importance=TRUE)

# Assess model outputs
rf_model6
rf_model7
rf_modelFull
sig_test<-rf.significance(rf_fit1000, rfselDF[2:length(rfselDF)])
imp_test<-importance(rf_fit[[1]])
plot(rf_fit500$rsq)
varImpPlot(rf_fit[[1]])
importance(rf_fit)
```

```{r}
predicted6<-rf_model6$predicted
predicted7<-rf_model7$predicted
predictedFull <-rf_modelFull$predicted

# plot observed values on the x-axis and predicted values on the y-axis. we are
# looking for these to be correlated (so close to a 1:1 line)
# Plot as Alpha ggplot
assess6 <- bind_cols(predict = predicted6, response = response)
assess7 <- bind_cols(predict = predicted7, response = response)
assessFull <- bind_cols(predict = predictedFull, response = response)

test<-matrix(c(response, predicted6, predicted7, predictedFull),nrow=len)

var6 <- text_grob(paste("Variance explained =", as.character(model6varEx[1:4]), sep = "\n"), x = .2, y = .9)
var7 <- text_grob(paste("Variance explained =", as.character(model7varEx[1:4]), sep = "\n"), x = .2, y = .9)
varFull <- text_grob(paste("Variance explained =", as.character(modelFullvarEx[1:4]), sep = "\n"), x = .2, y = .9)

gg6var <- ggplot(assess6, aes(x=response, y = predict)) +
  geom_point(alpha=.2, color = 'blue') +
  geom_smooth(method='lm',formula=y~x, color = 'red') +
  labs(x = "Observed diversity", y = "Predicted diversity") +
  annotation_custom(var6) +
  theme_few()
gg7var <- ggplot(assess7, aes(x=response, y = predict)) +
  geom_point(alpha=.2, color = 'blue') +
  geom_smooth(method='lm',formula=y~x, color = 'red') +
  labs(x = "Observed diversity", y = "Predicted diversity") +
  annotation_custom(var7) +
  theme_few()
ggfullvar <- ggplot(assessFull, aes(x=response, y = predict)) +
  geom_point(alpha=.2, color = 'blue') +
  geom_smooth(method='lm',formula=y~x, color = 'red') +
  labs(x = "Observed diversity", y = "Predicted diversity") +
  annotation_custom(varFull) +
  theme_few()
model_comp <- egg::ggarrange(gg6var, ggfullvar, nrow = 1)
annotate_figure(model_comp, top = text_grob("Comparing models with 6 and 14 variables", face = 'bold', size = 18))

# Calculate RMSE for the models
error6 <- predicted6 - response
error7 <- predicted7 - response
errorFull <- predictedFull - response

sqError6 <- error6^2
sqError7 <- error7^2
sqErrorFull <- errorFull^2

rmse <- function(error)
{
  sqrt(mean(error^2))
}

RMSE6 <- rmse(error6)
RMSE7 <- rmse(error7)
RMSEfull <- rmse(errorFull)

AIM_error6 <- bind_cols(error6 = error6,sqError6 = sqError6)
AIM_error7 <- bind_cols(error7 = error7,sqError7 = sqError7)
AIM_errorFull <- bind_cols(errorFull = errorFull,sqErrorFull = sqErrorFull)
AIM_error <- AIM_model %>%
  dplyr::select(1:2,32) %>%
  bind_cols(AIM_error6) %>%
  bind_cols(AIM_error7) %>%
  bind_cols(AIM_errorFull)

class(AIM_error)
View(AIM_error)
AIM_error_spatial <- st_sf(AIM_error)
HighError <- AIM_error_spatial %>%
  filter(errorFull >= 12 | errorFull <= -12) 
# 690 entries
RandomSample <- AIM_error[sample(1:nrow(AIM_error), nrow(HighError),
   replace=FALSE),]
RandomSampleSpat <-st_sf(RandomSample)

# Plot KDEs for each
highSP <- as(HighError, "Spatial")
highKDE <- sp.kde(testSP,n = 5000, bw = .4)
plot(highKDE)

writeRaster(highKDE,filename = 'data/Tiffs/highErrorKDE.tif', format = "GTiff")
st_write(HighError, 'data/Working/HighError.shp')

randomSP <- as(RandomSampleSpat, "Spatial")
randomKDE <- sp.kde(randomSP,n = 5000, bw = .4)
plot(randomKDE)

writeRaster(randomKDE,  filename = 'data/Tiffs/randomSampleKDE.tif', format = "GTiff")
st_write(RandomSample, 'data/Working/RandomSample.shp')

mapview(HighError,cex = 3, zcol = 'errorFull')

mapview(RandomSampleSpat,cex = 3, zcol = 'errorFull')
```



### Load GEE 2017 data for predictive mapping
```{r}
ndvi2017<- raster('data/Tiffs/GEE2017/ndvi2017.tif')
nbr2017 <- raster('data/Tiffs/GEE2017/nbr2017.tif')
tcb2017 <- raster('data/Tiffs/GEE2017/tcb2017.tif')
tcg2017 <- raster('data/Tiffs/GEE2017/tcg2017.tif')
tcw2017 <- raster('data/Tiffs/GEE2017/tcw2017.tif')
plot(ndvi2017)
plot(nbr2017)
plot(tcb2017)
plot(tcg2017)
plot(tcw2017)

# Import independent validation data
AIM_covariates <- read_sf('data/PlotsShapefiles/AIM_validate.shp') %>%
  st_transform(crs = 4326)
View(AIM_covariates)


testPredict <- predict(rf_fit1000)
### Reports accuracy, Cohen's kappa and area under ROC curve
testAssessPred <- accuracy(x = rf_fit1000, resp)
```

```{r}
# Import shapefile to clip
GBsubset <- read_sf('data/Shapefiles/ProjectSubset.shp') %>%
  st_transform(crs = 4326)

# Rasters needed to predict
tcb <- raster('data/Tiffs/GEE2017/tcb2017.tif')
tcg <- raster('data/Tiffs/GEE2017/tcg2017.tif')
cli_ffp <- raster('data/Tiffs/Cli/climate_ffp.tif')
trasp <- raster('data/Tiffs/Phys/GBtrasp.tif')
soil_h20 <- raster('data/Tiffs/Soil/SoilWaterMosaic.tif')

# Clip rasters to GB bounds
tcb_clip <- crop(tcb, GBsubset)
tcg_clip <- crop(tcg, GBsubset)
cli_ffp_clip <- crop(cli_ffp, GBsubset)
trasp_clip <- crop(trasp, GBsubset)
soil_h20_clip <- crop(soil_h20, GBsubset)
plot(tcb_clip)
plot(tcg_clip)
plot(cli_ffp_clip)
plot(trasp_clip)
plot(soil_h20_clip)

trasp_clip$extent
tcb_clip
tcg_clip
cli_ffp_clip
soil_h20_clip

mapview(GBsubset)

extent_proj = st_transform(extent,crs = 4326)
mapview(extent)

tcb_proj <- projectRaster(tcb_clip, crs="+proj=longlat +datum=WGS84")
tcg_proj <- projectRaster(tcg_clip, crs="+proj=longlat +datum=WGS84")
cli_ffp_proj <- projectRaster(cli_ffp_clip, crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
trasp_proj <- projectRaster(trasp_clip, crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
soil_h20_proj <- projectRaster(soil_h20_clip, crs="+proj=longlat +datum=WGS84")

stack_5band=stack(tcb_clip, tcg_clip, cli_ffp_clip, clip, clip)

latitude

```
