---
title: "Prep_MTBS_for_GEE"
author: "Eric Jensen"
date: "March 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
library(raster)
library(sf)
library(tidyverse)
library(lubridate)
```

Files that I want to keep are:  
- Prefire landsat image (possibly processed to keep only NIR and SWIR images)
- Postfire landsat image (possibly processed to keep only NIR and SWIR images)
- dNBR
- rdNBR
- Metadata file
Note: Need to somehow attach the year of the fire to the image

Delete files with phrases:
- burn_bndy
- gapmask
- mask
- desc
- map
- rep
- .kmz

Unzip files in directory
```{r}
# Set fire ids as file_str and setwd
wd <- "C:/Users/erjensen/Desktop/MTBS_AllData/"
annual_folders <- list.files(wd, full.names = T)
setwd(wd)
```

Produce function to keep only the dnbr and rdnbr images for all folders 1985-2017♫
```{r}
#Nested functions to retrieve only dnbr and rdnbr images
keep_dnbr_all <- function(year_folder){
  keep_dnbr <- function(Path){
    
    print(Path)
    print(list.files(Path))

    # Create list of files to delete
    files_to_delete <- list.files(path = Path, pattern =  'gapmask|mask|desc|map|rep|.kmz|refl|dnbr6|nbr5', full.names = T,recursive = T)
    print(files_to_delete)
  
    #delete files
    file.remove(files_to_delete)  }

  # Map the keep_dnbr function across all of the annual fire subfolders
  map(year_folder, keep_dnbr) }

map(annual_folders, keep_dnbr_all)
```


Get dnbr, rdnbr, and shapefiles and put them in separate folders
```{r}
#load package for moving files
library(filesstrings)

#create folders for outputs
# Function to grab and move RdNBR, dNBR, shapefiles, and metadata files to their own folders

move_sev_files = function(year_folder){
  dir.create(paste(year_folder, "/dNBR", sep=""))
  dir.create(paste(year_folder, "/RdNBR", sep=""))
  dir.create(paste(year_folder, "/shapefiles", sep=""))
  dir.create(paste(year_folder, "/metadata", sep=""))

  #Grab and move dnbr files
  dnbr_files <- list.files(path = year_folder, pattern = '_dnbr', recursive = TRUE, full.names = T)
  file.move(files = dnbr_files,destinations = paste(year_folder, "/dnbr", sep=""))

  #Grab and move rdnbr files
  rdnbr_files <- list.files(path = year_folder, pattern = '_rdnbr', recursive = TRUE, full.names = T)
  file.move(files = rdnbr_files,destinations = paste(year_folder, "/rdnbr", sep=""))

  #Grab and move shapefiles
  shapefiles <- list.files(path = year_folder, pattern = 'burn_bndy', recursive = TRUE, full.names = T)
  file.move(files = shapefiles,destinations = paste(year_folder, "/shapefiles", sep=""))
  
  #Grab and move metadata files
  metadata_files <- list.files(path = year_folder, pattern = 'metadata', recursive = TRUE, full.names = T)
  file.move(files = metadata_files,destinations = paste(year_folder, "/metadata", sep=""))

  # Delete all folders except dnbr and rdnbr folders
  remaining_folders <- list.files(path = year_folder, pattern = 'nv|az|ut|or|id|ca', full.names = T)
  unlink(remaining_folders,recursive=T)
}

#test the function
all_data <- "C:/Users/erjensen/Desktop/MTBS_AllData"
annual_folders <- list.files(all_data, full.names = T)
map(annual_folders, move_sev_files)
```

Clear the environment of variables
```{r}
rm(list = ls())
```


Because the fire severity data are extended out to the full extent of the image, rather than masked at the fire boundary, this function masks each fire's dNBR and RdNBR file by the corresponding shapefile with which is was packaged.
```{r}
annual_lists <- list.files("C:/Users/erjensen/Desktop/MTBS_AllData/", full.names = T)

# nested function to mask each RdNBR and dNBR raster by the boundary of the fire and then export as a new tiff
annual_fires <- function(year_folder){
    
    # Get list of fire ids for entire year to filter dnbr, rdnbr, and shapefiles by
    fire_list <- list.files(paste(year_folder,'/metadata', sep=''))
    print(fire_list)
    fire_ids <- fire_list %>%
                    map(substr, start = 1, stop = 21) %>%
                    unlist() %>%
                    map(str_to_lower) %>%
                    unlist()
    
    # Get list of dnbr, rdnbr, and shapefiles
    dnbr_list <- list.files(paste(year_folder, '/dNBR', sep= ''), pattern = '.tif', full.names = T)
    rdnbr_list <- list.files(paste(year_folder, '/RdNBR', sep = ''), pattern = '.tif', full.names = T)
    shp_list <- list.files(paste(year_folder, '/shapefiles', sep = ''), pattern = '.shp', full.names = T)
    
    # Create new directories to store clipped rasters once written
    dir.create(paste(year_folder, "/dNBR_masked", sep=""))
    dir.create(paste(year_folder, "/RdNBR_masked", sep=""))
    
    mask_fires <- function(fire_id){

        # Subset lists of files by the fire id
        dnbr_file <- str_subset(dnbr_list,fire_id)
        print(dnbr_file)
        #print(length(dnbr_file))
        rdnbr_file <- str_subset(rdnbr_list,fire_id)
        #print(length(rdnbr_file))
        shp_file <- str_subset(shp_list,fire_id)
        #print(length(shp_file))
    
        # If statement to skip if there are not dnbr files
        if(length(dnbr_file) != 0){
        
          # import spatial files for masking as raster and sf
          shp_sf <- st_read(shp_file, quiet=T)
          dnbr_rast <- raster(dnbr_file)
          rdnbr_rast <- raster(rdnbr_file)
          
          # Mask the dnbr and rdnbr files by the sf object
          dnbr_masked <- mask(dnbr_rast, shp_sf)
          rdnbr_masked <- mask(rdnbr_rast, shp_sf)
      
          # Write new dnbr and rdnbr rasters to the appropriate folder with an appropriate name
          writeRaster(dnbr_masked, paste(year_folder, '/dNBR_masked/', substr(dnbr_file,50,89), 'dnbrmasked.tif', sep = ''), overwrite = T)
          writeRaster(rdnbr_masked, paste(year_folder, '/rdNBR_masked/', substr(dnbr_file,50,89), 'rdnbrmasked.tif', sep = ''), overwrite = T) 
          }
        else {}
      }
    map(fire_ids, mask_fires)
    }

# Apply nested function to entire folder stack
map(annual_lists, annual_fires)
```


Import and mosaic masked rasters for each year
```{r}
# list of annual folder paths
annual_lists <- list.files("C:/Users/erjensen/Desktop/MTBS_AllData/", full.names = T)

# Function to produce annual mosaics
annual_mosaics <- function(year_folder){
  
    # Get year string for parsing output filename
    year <- substr(year_folder, start = 40, stop = 43)
    print(year)

    # Get filelists for masked dnbr and rasters
    dnbr_list <- list.files(paste(year_folder, '/dNBR_masked', sep= ''), pattern = 'dnbrmasked.tif$', full.names = T)
    print(dnbr_list)
    rdnbr_list <- list.files(paste(year_folder, '/RdNBR_masked', sep = ''), pattern = 'rdnbrmasked.tif$', full.names = T)
    print(rdnbr_list)
    
    # Import masked dnbr and rdnbr rasters
    dnbr_rasters <- map(dnbr_list,raster)
    rdnbr_rasters <- map(rdnbr_list,raster)
    
    # Mosaic dnbr and rdnbr files
    dnbr_rasters$fun <- mean # set function
    dnbr_mosaic <- do.call(mosaic, dnbr_rasters)
    rdnbr_rasters$fun <- mean # set function
    rdnbr_mosaic <- do.call(mosaic, rdnbr_rasters)
    
    # Write the raster mosaics to be imported into GEE
    writeRaster(dnbr_mosaic, filename = paste(year_folder, '/dNBRmosaic', year, '.tif', sep = ''))
      print(paste(year_folder, '/dNBRmosaic', year, '.tif', sep = ''))
    writeRaster(rdnbr_mosaic, filename = paste(year_folder, '/RdNBRmosaic', year, '.tif', sep = ''))  
      print(paste(year_folder, '/RdNBRmosaic', year, '.tif', sep = ''))
                }

map(annual_lists, annual_mosaics)
```
