---
title: "ProcessGEEZonalStats"
author: "Eric Jensen"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(mapview)
library(reshape2)
```

### Import zonal statistics CSVs and process to prepare for join
```{r}
# list the csv files for each zonal statistic
ndvi_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNDVI", full.names = T)
ndvi_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNDVI", full.names = T)
nbr_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanNBR", full.names = T)
nbr_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdNBR", full.names = T)
tcb_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCB", full.names = T)
tcb_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCB", full.names = T)
tcg_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCG", full.names = T)
tcg_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCG", full.names = T)
tcw_mean_zonal <- list.files(path = "data/zonalCSVs", pattern = "meanTCW", full.names = T)
tcw_std_zonal <- list.files(path = "data/zonalCSVs", pattern = "stdTCW", full.names = T)
listofLists = list(ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

RScovariates <- length(listofLists)

# create empty lists for each statistic
ndvi_mean <- list()
ndvi_std <- list()
nbr_mean <- list()
nbr_std <- list()
tcb_mean <- list()
tcb_std <- list()
tcg_mean <- list()
tcg_std <- list()
tcw_mean <- list()
tcw_std <- list()

# functions to read the CSVs into the list
ndvi_mean <- map(ndvi_mean_zonal,read_csv)
ndvi_std <- map(ndvi_std_zonal,read_csv)
nbr_mean <- map(nbr_mean_zonal,read_csv)
nbr_std <- map(nbr_std_zonal,read_csv)
tcb_mean <- map(tcb_mean_zonal,read_csv)
tcb_std <- map(tcb_std_zonal,read_csv)
tcg_mean <- map(tcg_mean_zonal,read_csv)
tcg_std <- map(tcg_std_zonal,read_csv)
tcw_mean <- map(tcw_mean_zonal,read_csv)
tcw_std <- map(tcw_std_zonal,read_csv)
remove(listofLists,ndvi_mean_zonal,ndvi_std_zonal,nbr_mean_zonal,nbr_std_zonal,tcb_mean_zonal,
                   tcb_std_zonal,tcg_mean_zonal,tcg_std_zonal,tcw_mean_zonal,tcw_std_zonal)

#bind cols to produce dataframes of each zonal statistic, rename cols with unique name
ndvi_mean <- ndvi_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_mean = mean)
ndvi_std <- ndvi_std %>% bind_rows() %>% dplyr::select(PrimaryKey, ndvi_stdDev = stdDev)
nbr_mean <- nbr_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_mean = mean)
nbr_std <- nbr_std %>% bind_rows() %>% dplyr::select(PrimaryKey, nbr_stdDev = stdDev)
tcb_mean <- tcb_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_mean = mean)
tcb_std <- tcb_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcb_stdDev = stdDev)
tcg_mean <- tcg_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_mean = mean)
tcg_std <- tcg_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcg_stdDev = stdDev)
tcw_mean <- tcw_mean %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_mean = mean)
tcw_std <- tcw_std %>% bind_rows() %>% dplyr::select(PrimaryKey, tcw_stdDev = stdDev)

# Assessed all dataframes for missing predictor variable values, none present
```

# Import spatial plots and join with GEE zonal stats
```{r}
# Import plots sf for joining GEE zonal stats too
AIM_plots <- read_sf('data/AIM_model.shp') %>%
  st_transform(crs = 4326)
# Convert to tibble without geometry for manipulating and joining
AIMdf <- as_data_frame(matrix(unlist(AIM_plots), nrow=nrow(AIM_plots), byrow=F))
AIM_plotsDF <- AIMdf %>%
  dplyr::select(-V3, -V7, -V8) %>%
  dplyr::rename( "plotkey" = V1, "PlotID" = V2, "VisitYear" = V4, 
                 "PrimaryKey" = V5, "SpeciesLis" = V6)

listofGEE = list(AIM_plotsDF,ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std)
# Join the zonal statistics to each plot, remove plots with NAs--likely only the dateless plot
AIM_wGEE <- reduce(listofGEE, left_join)
remove(ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std, listofGEE, AIMdf)
```

### Calculate zonal statistics for rasters of elevation, slope, aspect, etc.
```{r}
######################################################
######## Change to buffers once up and running #######
######################################################
AIM_buffers <- read_sf('data/AIM_modelbuff_100m.shp') %>%
  st_transform(crs = 4326)

# bind primary key to join by
AIM_wGEE <- AIM_wGEE %>%
  left_join(AIM_plots, "PrimaryKey") %>%
  dplyr::select(-ends_with(".y"))

# subset remote sensing covariate columns to bind to sf object
to_join<- as_tibble(AIM_wGEE[6:(5+RScovariates)])
AIM_RS = bind_cols(AIM_plots,to_join)
View(AIM_RS)

#####################################################
########## Check aspect, HLI? #######################
#####################################################

# import physical rasters
GBdem<- raster('data/Tiffs/GBdem_new.tif')
GBslope<- raster('data/Tiffs/GBslope.tif')
GBaspect<- raster('data/Tiffs/GBaspect.tif')
GBtpi<- raster('data/Tiffs/GBtpi.tif')

# extract raster values and mutate new column to prediction layers
AIM_RS_Phys <- as_tibble(AIM_RS) %>%
  dplyr::mutate(elev = raster::extract(x=GBdem, y=AIM_RS)) %>%
  dplyr::mutate(aspect = raster::extract(x=GBaspect, y=AIM_RS)) %>%
  dplyr::mutate(slope = raster::extract(x=GBslope, y=AIM_RS)) %>%
  dplyr::mutate(tpi = raster::extract(x=GBtpi, y=AIM_RS))

AIM_RS_Phys <-AIM_wGEE_wPhys
View(AIM_RS_Phys)

# extract
```

### Calculate species richness values based on list
```{r}
# Parse the species list strings to lists
AIM_RS_Phys_SR <- AIM_RS_Phys %>%
  dplyr::mutate(SpecList = str_split(SpeciesLis,pattern=";")) %>%
  dplyr::select(-SpeciesLis)
View(AIM_RS_Phys_SR)

# Calculate number of species at each plot (-1 because of extra "") and bind the dataframe to predictive dataframe
SpecN = list()
for (i in 1:nrow(AIM_RS_Phys_SR)) {
          SpecN[i] <- length(AIM_RS_Phys_SR$SpecList[[i]]) - 1}
SpecNdf <- data.frame(matrix(unlist(SpecN), nrow=length(SpecN), byrow=T))
AIM_RS_Phys_SR_count <- bind_cols(AIM_RS_Phys_SR,SpecNdf)

View(AIM_RS_Phys_SR_count)
AIM_PhysRS <- AIM_wGEE_wPhys_wSR
remove(AIM_wGEE,AIM_wGEE_wPhys,AIM_wGEE_wPhys_wSR)

# Write existing table to CSV
st_write(AIM_RS_Phys_SR_count, 'data/Working/AIM_covariates_wrk.shp')

# Read in USDA plants dataframe to get genus and species data
# plantsList <- read_csv('data/TXT/USDA_PlantsTable.txt')
```


### Read-in output CSV above to avoid running everything upstream
```{r}
AIM_working <- read_sf('data/Working/AIM_covariates_wrk.shp') %>%
  st_transform(crs = 4326)
# View(AIM_working)
# mapview(AIM_working)
```


### Assess climate covariate data (36 deg - 45 deg; 112 deg - 122 deg)
#### Climate variables of interest are:
- ADI (annual dryness index)
- SDI (summer dryness index)
- FFP (frost free period)
- GSP (growing season precipitation)
- MAP (mean annual precipitation)
- SPRP (spring precipitation)
- SMRP (summer precipitation)
```{r}
### Import CSV from previous lines

### Import mosaics of desired variables
cliIndexList <- c("_adi.tif", "_sdi.tif", "_sprp.tif", "_smrp.tif", "_ffp.tif", "_gsp.tif", "_map.tif")
cliFolderList <- list()
cliFolderList <- grep(list.files(path = 'D:/current_climates', full.names = T), pattern = "33|34|35|111|123|124", invert = T, value = T)

### Only create the list once--otherwise it will overwrite
cliList <- list() 
cliMosaic <- list()

### Vary the index from 1:the number of indices
index = 1
cliIndexList[[index]]
cliList[[index]] <- list()
# cliList[[index]]$index <- str_replace(cliIndexList[[index]], ".tif", "")
# cliList[[index]]$index <- str_replace(cliList[[index]]$index, "_", "")

### Add each image into a list of images for a given index
for(i in seq_along(cliFolderList)){ 
    cliList[[index]][[i]] <- raster(grep(list.files(path = cliFolderList[[i]], full.names = T), 
                            pattern = cliIndexList[[index]], value = T))}
#View(cliList)
cliList[[index]]$fun <- mean


#-------------------------------------------------------------------------
### Some indices are missing some images--this code runs for those indices
parse1 = paste(cliFolderList, '/', sep = "")
parse2 = paste(parse1,(substr(cliFolderList,21,31)), sep = "")
adilist <- paste(parse2, cliIndexList[index], sep = "")
print(adilist)

adiraster <- list()
for(i in seq_along(adilist)){
  if (file.exists(adilist[i]))
    adiraster[i] <- raster(adilist[i])
}

for(i in 1:99){
  print(typeof(adiraster[[i]]))
}

adiraster$fun <- mean
adiMosaic <- do.call(mosaic,adiraster)

names(cliMosaic[[index]]) <- cliIndexList[[index]] %>% str_replace(".tif", "") %>% str_replace("_", "")

#-------------------------------------------------------------------------
### Mosaic images together for each index
cliList[[index]]$fun <- mean

cliMosaic[[index]] <- do.call(mosaic,cliList[[index]])
names(cliMosaic[[index]]) <- cliIndexList[[index]] %>% str_replace(".tif", "") %>% str_replace("_", "")

View(cliMosaic[[index]])
plot(cliMosaic[[index]])

#-------------------------------------------------------------------------
### Write rasters

writeRaster(cliMosaic[[1]], filename = 'data/Tiffs/climate_adi.tif', format = "GTiff")
writeRaster(cliMosaic[[2]], filename = 'data/Tiffs/climate_sdi.tif', format = "GTiff")
# writeRaster(cliMosaic[[3]], filename = 'data/Tiffs/climate_sprp.tif', format = "GTiff")
# writeRaster(cliMosaic[[4]], filename = 'data/Tiffs/climate_smrp.tif', format = "GTiff")
# writeRaster(cliMosaic[[5]], filename = 'data/Tiffs/climate_ffp.tif', format = "GTiff")
# writeRaster(cliMosaic[[6]], filename = 'data/Tiffs/climate_gsp.tif', format = "GTiff")
# writeRaster(cliMosaic[[7]], filename = 'data/Tiffs/climate_map.tif', format = "GTiff")

Climate_adi <- raster('data/Tiffs/climate_adi.tif')
Climate_sdi <- raster('data/Tiffs/climate_sdi.tif')
Climate_sprp <- raster('data/Tiffs/climate_sprp.tif')
Climate_smrp <- raster('data/Tiffs/climate_smrp.tif')
Climate_ffp <- raster('data/Tiffs/climate_ffp.tif')
Climate_gsp <- raster('data/Tiffs/climate_gsp.tif')
Climate_map <- raster('data/Tiffs/climate_map.tif')

plot(Climate_map)
```

### Calculate zonal statistics for each climate variable of interest
```{r}

```



### Import soils data
```{r}
SoilWater <- list()
SoilWater[[1]] <- raster('data/Tiffs/Soil/SoilWater1.tif')
SoilWater[[2]] <- raster('data/Tiffs/Soil/SoilWater2.tif')
SoilWater[[3]] <- raster('data/Tiffs/Soil/SoilWater3.tif')
SoilWater[[4]] <- raster('data/Tiffs/Soil/SoilWater4.tif')
SoilWater$fun <- mean
SoilWaterMosaic <- do.call(mosaic,SoilWater)

SoilCarbon <- list()
SoilCarbon[[1]] <- raster('data/Tiffs/Soil/Carbon1.tif')
SoilCarbon[[2]] <- raster('data/Tiffs/Soil/Carbon2.tif')
SoilCarbon[[3]] <- raster('data/Tiffs/Soil/Carbon3.tif')
SoilCarbon[[4]] <- raster('data/Tiffs/Soil/Carbon4.tif')
SoilCarbon$fun <- mean
SoilCarbonMosaic <- do.call(mosaic,SoilCarbon)

SoilTexture <- list()
SoilTexture[[1]] <- raster('data/Tiffs/Soil/TextureClass1.tif')
SoilTexture[[2]] <- raster('data/Tiffs/Soil/TextureClass2.tif')
SoilTexture[[3]] <- raster('data/Tiffs/Soil/TextureClass3.tif')
SoilTexture[[4]] <- raster('data/Tiffs/Soil/TextureClass4.tif')
SoilTexture$fun <- mean
SoilTextureMosaic <- do.call(mosaic,SoilTexture)

SoilSand <- list()
SoilSand[[1]] <- raster('data/Tiffs/Soil/Clay1.tif')
SoilSand[[2]] <- raster('data/Tiffs/Soil/Clay2.tif')
SoilSand[[3]] <- raster('data/Tiffs/Soil/Clay3.tif')
SoilSand$fun <- mean
SoilSandMosaic <- do.call(mosaic,SoilSand)

SoilClay <- list()
SoilClay[[1]] <- raster('data/Tiffs/Soil/Sand1.tif')
SoilClay[[2]] <- raster('data/Tiffs/Soil/Sand2.tif')
SoilClay[[3]] <- raster('data/Tiffs/Soil/Sand3.tif')
SoilClay[[4]] <- raster('data/Tiffs/Soil/Sand4.tif')
SoilClay$fun <- mean
SoilClayMosaic <- do.call(mosaic,SoilClay)

writeRaster(SoilWaterMosaic, filename = 'data/Tiffs/Soil/SoilWaterMosaic.tif', format = "GTiff")
writeRaster(SoilCarbonMosaic, filename = 'data/Tiffs/Soil/SoilCarbonMosaic.tif', format = "GTiff")
writeRaster(SoilTextureMosaic, filename = 'data/Tiffs/Soil/SoilTextureMosaic.tif', format = "GTiff")
writeRaster(SoilSandMosaic, filename = 'data/Tiffs/Soil/SoilSandMosaic.tif', format = "GTiff")
writeRaster(SoilClayMosaic, filename = 'data/Tiffs/Soil/SoilClayMosaic.tif', format = "GTiff")

```

### Calculate zonal statistics for each soils variable of interest
```{r}

```







