---
title: "RandomForest 2.0"
author: "Eric Jensen"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(randomForest)
library(rfUtilities)
library(corrplot)
library(ggthemes)
library(egg)
library(ggpubr)
library(spatstat)
```

### Import zonal statistics CSVs and process to prepare for join
```{r}
BioPhys_mean = read_csv('data/GEEZonal_CSVs/Biophysical_Predictors_Mean.csv') %>% drop_na()
BioPhys_mode = read_csv('data/GEEZonal_CSVs/Biophysical_Predictors_Mode.csv') %>% drop_na()
Climate_ann = read_csv('data/GEEZonal_CSVs/Climate_covariates.csv') %>% drop_na()
Climate_mean = read_csv('data/GEEZonal_CSVs/Climate_Predictors_Mean.csv') %>% drop_na()
Soil = read_csv('data/GEEZonal_CSVs/Soil_Predictors.csv') %>% drop_na()
Fire = read_csv('data/GEEZonal_CSVs/Fire_Predictors.csv')
RemSens = read_csv('data/GEEZonal_CSVs/RS_Predictors.csv') %>% drop_na()
RemSens_clust = read_csv('data/GEEZonal_CSVs/RS_clust_predictors.csv') %>% drop_na()
```

Remove NAs from remote sensing data
```{r}
RemSens_clean <- RemSens %>%
  dplyr::select(-c(.geo, VisitYear)) %>%
  dplyr::select(-1) %>%
  drop_na() %>%
  mutate(B1_cv = B1_sd/B1_mn) %>% #Mutate coefficent of variation columns
  mutate(B2_cv = B2_sd/B2_mn) %>%
  mutate(B3_cv = B3_sd/B3_mn) %>%
  mutate(B4_cv = B4_sd/B4_mn) %>%
  mutate(B5_cv = B5_sd/B5_mn) %>%
  mutate(B7_cv = B7_sd/B7_mn) %>%
  mutate(MSAVI2_cv = MSAVI2_sd/MSAVI2_mn) %>%
  mutate(SAVI_cv = SAVI_sd/SAVI_mn) %>%
  mutate(SATVI_cv = SATVI_sd/SATVI_mn) %>%
  mutate(NBR_cv = NBR_sd/NBR_mn) %>%
  mutate(NDMI_cv = NDMI_sd/NDMI_mn) %>%
  mutate(NDVI_cv = NDVI_sd/NDVI_mn) %>%
  mutate(TCA_cv = TCA_sd/TCA_mn) %>%
  mutate(TCG_cv = TCG_sd/TCB_mn) %>%
  mutate(TCW_cv = TCW_sd/TCW_mn) %>%
  mutate(TCB_cv = TCB_sd/TCB_mn) 
```

Calculate proportion of distinct classes to number of cells in unsupervised classification data
```{r}
RemSens_clust = RemSens_clust %>%
  mutate(k25_prop = kmeans25_distcount/kmeans25_count) %>%
  mutate(k50_prop = kmeans50_distcount/kmeans50_count) %>%
  mutate(k100_prop = kmeans100_distcount/kmeans100_count) %>%
  mutate(k200_prop = kmeans200_distcount/kmeans200_count) %>%
  select(-c('kmeans25_count','kmeans50_count','kmeans100_count','kmeans200_count','CountNorm', contains('histogram'))) %>%
  filter(k25_prop != 'inf')
```

Clean the imported predictor CSVs
```{r}
BioPhys_mean_clean = BioPhys_mean %>%
  dplyr::select(-c(PlotID, Program, VisitYear, .geo, SpeciesN, VisitDate)) %>%
  dplyr::select(-1) 

clean_df <- function(df){
  df %>%
  dplyr::select(-c(PlotID, Program, .geo, SpeciesN, VisitDate, VisitYear,1))}

BioPhys_mode_clean <- clean_df(BioPhys_mode) %>% mutate(landform = mode) %>% dplyr::select(-mode)
Climate_ann_clean <- clean_df(Climate_ann)
Climate_mean_clean <- clean_df(Climate_mean)
Soil_clean <- clean_df(Soil)
RemSens_clust_clean <- clean_df(RemSens_clust) %>% dplyr::select(-c(contains('_count'), contains('prop')))
Fire_clean <- clean_df(Fire) %>% dplyr::select(-c(YearsOfMTBS,LastFire))%>%replace(., is.na(.), 0)

remove(BioPhys_mean,BioPhys_mode,Climate_ann,Climate_mean,Soil,Fire,RemSens,RemSens_clust, clean_df)
```

Join the dataframes together in multiple ways
ALL PREDICTORS
```{r}
Predictors_all <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  left_join(Climate_mean_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  dplyr::rename(PrimaryKey = p.PrimaryKey) # revert names of non-predictors

#not sure where the 70 nas are coming from

#remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean)
```

REMOVING PREDICTORS FOR INDIVIDUAL BANDS
```{r}
Predictors_noBands <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean%>%select(-starts_with('B')), by = 'PrimaryKey') %>% #remove columns that start with 'B'
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  left_join(Climate_mean_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  dplyr::rename('PrimaryKey' = 'p.PrimaryKey') # revert names of non-predictors
  
  #not sure where the 70 nas are coming from

remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean)
```

# Import spatial plots and join with GEE zonal stats--largely unnecessary
```{r}
# library(sf)
# # Import plots sf for joining GEE zonal stats too
# AIM_plots <- read_sf('data/Shapefiles/AIM_model.shp') %>%
#   st_transform(crs = 4326)
# # Convert to tibble without geometry for manipulating and joining
# AIMdf <- as_data_frame(matrix(unlist(AIM_plots), nrow=nrow(AIM_plots), byrow=F))
# AIM_plotsDF <- AIMdf %>%
#   dplyr::select(-V3, -V7, -V8) %>%
#   dplyr::rename( "plotkey" = V1, "PlotID" = V2, "VisitYear" = V4, 
#                  "PrimaryKey" = V5, "SpeciesLis" = V6)
# 
# listofGEE = list(AIM_plotsDF,ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std)
# # Join the zonal statistics to each plot, remove plots with NAs--likely only the dateless plot
# AIM_wGEE <- reduce(listofGEE, left_join)
```

### Import Species Diversity data
```{r}
Responses = read_csv('data/SpectralDiversity_CSVs/Allmodel.csv') %>% select(c(PrimaryKey,SpeciesN))
```

Join the predictors and responses together
```{r}
BuildModelTable = function(preds){
  table <- preds %>%
            full_join(Responses, 'PrimaryKey') %>%
            drop_na() 
  return(table)}

# All responses
ModelTable_all = BuildModelTable(Predictors_all)
# All predictors except individual bands
ModelTable_noBands = BuildModelTable(Predictors_noBands)

remove(Predictors_all,Predictors_noBands)
```

################## Random Forest Modeling ######################
### Prepare Data for Model Selection
```{r}
## Subset dataframe into predictors and responses
subset_preds = function(ModelTable){
       predictors = ModelTable %>%
                       dplyr::select(starts_with('p.')) %>%
                       mutate(p.landform_v = factor(ModelTable$p.landform)) %>%
                       dplyr::select(-p.landform)
       return(predictors)}

Preds_All = subset_preds(ModelTable_all)
Preds_NoBands = subset_preds(ModelTable_noBands)

subset_resp = function(ModelTable){
       responses = ModelTable %>%
                       dplyr::select(starts_with('Species'))
       return(responses)}

Resps_All = subset_resp(ModelTable_all)
remove(ModelTable_all,ModelTable_noBands,Responses)
```

```{r}
################################################################################
########################### Skipping this for now ##############################

# # Get mean correlation across all responses
# # Convert df to absolute values
# Corr_df_abs <- abs(Cor_df[2:20])
# predNames <- data.frame(Cor_df[1])
# predCor <- data.frame(rowMeans(Corr_df_abs))
# predRank <- cbind(predNames,predCor)
# View(predRank)
# 
# # Get mean correlation across all predictors
# colMeans(Corr_df_abs)
```

Random forest variable selection
```{r}
################################################################################
########################### Skipping this for now ##############################

RFsel_All_SR = rf.modelSel(xdata = Preds_All, ydata = Resps_All$SpeciesN, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
RFsel_NoBands_SR = rf.modelSel(xdata = Preds_NoBands, ydata = Resps_All$SpeciesN, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 

#Assess outputs from species richness model
plot(RFsel_All_SR)
plot(RFsel_NoBands_SR)
```

```{r}
################################################################################
########################### Skipping this for now ##############################

# Function to process the rf.modelSel() objects
Process_ModelSel = function(rfsel){
    rfsel <- cbind(rownames(rfsel$sel.importance), rfsel$sel.importance)
    rownames(rfsel) <- NULL
    colnames(rfsel) <- c("name", "imp")
    return(rfsel)}

RFsel_ALL_SR_ready = Process_ModelSel(RFsel_All_SR)
RFsel_NoBands_SR_ready = Process_ModelSel(RFsel_NoBands_SR)

###########################################################
######## Convert these to more succinct functions #########

# Order the dataframe by variable importance
RFsel_ALL_SR_ready <- RFsel_ALL_SR_ready[order(-RFsel_ALL_SR_ready$imp),]
RFsel_NoBands_SR_ready <- RFsel_NoBands_SR_ready[order(-RFsel_NoBands_SR_ready$imp),]

# Get list of names of top 30 predictors for each model
RFsel_ALL_SR_names <- RFsel_ALL_SR_ready$name[1:nrow(RFsel_ALL_SR_ready)]
RFsel_NoBands_SR_names <- RFsel_NoBands_SR_ready$name[1:nrow(RFsel_NoBands_SR_ready)]

# drop columns that are not as important based on rfmodelsel--columns not in list of names
RFsel_ALL_SR_cov <- Preds_All[,as.character(RFsel_ALL_SR_names)] #[-c(4,5,6,9,8,7)]]
RFsel_NoBands_SR_cov <- Preds_NoBands[,as.character(RFsel_NoBands_SR_names)]

#Too many GD object--get rid of some
remove(RFsel_ALL_SR_ready,RFsel_NoBands_SR_ready,Predictors_all,Predictors_noBands,BuildModelTable,Process_ModelSel,subset_preds,subset_resp)
```

Assess correlation between model predictors
```{r}
################################################################################
########################### Skipping this for now ##############################

################################################################################
########### Figure out better way to eliminate correlated variables ############

# calculate correlation coefficient matrix
# Figure out how to include landform
cor_All <-cor(RFsel_ALL_SR_cov%>%select(-p.landform_v), method="pearson")
cor_NoBands <-cor(RFsel_NoBands_SR_cov, method="pearson")

# plot the correlation. the darker the number, the more correlated the two
# variables
corrplot(cor_All,method="number")
corrplot(cor_NoBands,method="number")

## You can now manually remove any variables that are above your correlation 
## threshold. In this case we will just leave them all in, but we could remove
## them by using the following code

###### Model fitting
ALL_SR_model <- cbind(RFsel_ALL_SR_cov, specN= Resps_All$SpeciesN)
NoBands_SR_model <- cbind(RFsel_NoBands_SR_cov, specN= Resps_All$SpeciesN)

# Now let's restrict our final predictors to those that are most important (Change manually) 
# and run the continuous model
```

Run random forest models
```{r}
### Generate random number to set seed
# sample(1:1000000,1)
rf_seed = 517953
set.seed(rf_seed)

# Prepare dataframe for modeling
All_SR_model <- cbind(specN= Resps_All$SpeciesN, Preds_All)
NoBands_SR_model <- cbind(specN= Resps_All$SpeciesN, Preds_NoBands)

# Run tuning function
rf_all_tune <-tuneRF(y=Resps_All$SpeciesN, x = Preds_All, ntreeTry=300)
rf_nobands_tune <-tuneRF(y=Resps_All$SpeciesN, x = Preds_NoBands, ntreeTry=300)
print(rf_all_tune)

# Get the mtry value that minimizes error
mtry_all_opt <- rf_all_tune[, 'mtry'][which.min(rf_all_tune[,'OOBError'])]
mtry_nobands_opt <- rf_nobands_tune[, 'mtry'][which.min(rf_nobands_tune[,'OOBError'])]

# Apply the model using the optimal mtry tuning parameter
ALL_rf_model = randomForest(specN ~ ., data=All_SR_model, importance = TRUE, ntree = 200, mtry = mtry_all_opt)
NoBands_rf_model = randomForest(specN ~ ., data=NoBands_SR_model, importance = TRUE, ntree = 200, mtry = mtry_nobands_opt)

varImpPlot(ALL_rf_model)
varImpPlot(NoBands_rf_model)

# Use the strongest model going forward
NoBands_rf_model
ALL_rf_model
plot(NoBands_rf_model)
```

**Tuning a Random Forest via tree depth**  
From the Data Camp *Machine learning with decision trees* course
```{r}
# Establish a list of possible values for mtry, nodesize and sampsize
mtry <- seq(4, ncol(NoBands_SR_model) * 0.8, 12)
nodesize <- seq(3, 8, 2)
sampsize <- nrow(NoBands_SR_model) * c(0.7, 0.8)

# Create a data frame containing all combinations 
hyper_grid <- expand.grid(mtry = mtry, nodesize = nodesize, sampsize = sampsize)

# Create an empty vector to store OOB error values
rsq_err <- c()

# Write a loop over the rows of hyper_grid to train the grid of models
for (i in 1:nrow(hyper_grid)) {
    print(i)
    # Train a Random Forest model
    model <- randomForest(formula = specN ~ ., 
                          data = NoBands_SR_model,
                          mtry = hyper_grid$mtry[i],
                          nodesize = hyper_grid$nodesize[i],
                          sampsize = hyper_grid$sampsize[i],
                          ntree = 200)
                          
    # Store OOB error for the model  
    rsq_err[i] <- model$rsq[length(model$rsq)]
    print(rsq_err[i])
}

# Identify optimal set of hyperparmeters based on OOB error
opt_i <- which.max(rsq_err)
print(hyper_grid[opt_i,])
```


Plot the predicted data vs. the observed data
```{r}
# Round for later plotting
NoBands_rf_modelEx <- round(NoBands_rf_model$rsq[3000], digits = 4)

NoBands_predicted <- NoBands_rf_model$predicted

# plot observed values on the x-axis and predicted values on the y-axis. we are
# looking for these to be correlated (so close to a 1:1 line)
# Plot as Alpha ggplot
NoBands_assess <- bind_cols(predict = NoBands_predicted, response = Resps_All$SpeciesN)

varNoBands <- text_grob(paste("Variance explained =", as.character(NoBands_rf_modelEx[1:4]), sep = "\n"), x = .2, y = .9)

gg_NoBands <- ggplot(NoBands_assess, aes(x=response, y = predict)) +
  geom_point(alpha=.1, color = 'blue',  shape = 16) +
  geom_smooth(method='lm',formula=y~x, color = 'red') +
  labs(x = "Observed diversity", y = "Predicted diversity",title="28 variables, 1000 trees") +
  annotation_custom(varNoBands) +
  theme_few()
gg_NoBands

# calculate r-squared and RMSE
actual = NoBands_assess$response
predicted = NoBands_assess$predict
R2 <- 1 - (sum((actual-predicted)^2)/sum((actual-mean(actual))^2))
library(caret)
rmse <- RMSE(NoBands_assess$predict,NoBands_assess$response)

# model_comp <- egg::ggarrange(gg1000var,gg3000var,gg5000var, nrow = 1)
# annotate_figure(model_comp, top = text_grob("Comparing models with 6 and 14 variables", face = 'bold', size = 18))
```

