---
title: "ProcessGEEZonalStats"
author: "Eric Jensen"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(randomForest)
library(rfUtilities)
library(corrplot)
library(ggthemes)
library(egg)
library(ggpubr)
library(spatstat)
```

### Import zonal statistics CSVs and process to prepare for join
```{r}
BioPhys_mean = read_csv('data/GEEZonalCSVs/Biophysical_Predictors_Mean.csv')
BioPhys_mode = read_csv('data/GEEZonalCSVs/Biophysical_Predictors_Mode.csv')
Climate_ann = read_csv('data/GEEZonalCSVs/Climate_predictors_annual.csv')
Climate_mean = read_csv('data/GEEZonalCSVs/Climate_Predictors_Mean.csv')
Soil = read_csv('data/GEEZonalCSVs/Soil_Predictors.csv')
Fire = read_csv('data/GEEZonalCSVs/Fire_Predictors.csv')
RemSens = read_csv('data/GEEZonalCSVs/RS_Predictors.csv')
RemSens_clust = read_csv('data/GEEZonalCSVs/RS_clust_predictors.csv')
```

Remove NAs from remote sensing data
```{r}
RemSens_clean <- RemSens %>%
  dplyr::select(-c(.geo, VisitYear)) %>%
  dplyr::select(-1) %>%
  drop_na() %>%
  mutate(B1_cv = B1_sd/B1_mn) %>% #Mutate coefficent of variation columns
  mutate(B2_cv = B2_sd/B2_mn) %>%
  mutate(B3_cv = B3_sd/B3_mn) %>%
  mutate(B4_cv = B4_sd/B4_mn) %>%
  mutate(B5_cv = B5_sd/B5_mn) %>%
  mutate(B7_cv = B7_sd/B7_mn) %>%
  mutate(MSAVI2_cv = MSAVI2_sd/MSAVI2_mn) %>%
  mutate(SAVI_cv = SAVI_sd/SAVI_mn) %>%
  mutate(SATVI_cv = SATVI_sd/SATVI_mn) %>%
  mutate(NBR_cv = NBR_sd/NBR_mn) %>%
  mutate(NDMI_cv = NDMI_sd/NDMI_mn) %>%
  mutate(NDVI_cv = NDVI_sd/NDVI_mn) %>%
  mutate(TCA_cv = TCA_sd/TCA_mn) %>%
  mutate(TCG_cv = TCG_sd/TCB_mn) %>%
  mutate(TCW_cv = TCW_sd/TCW_mn) %>%
  mutate(TCB_cv = TCB_sd/TCB_mn) 
```

Calculate proportion of distinct classes to number of cells in unsupervised classification data
```{r}
RemSens_clust = RemSens_clust %>%
  mutate(k25_prop = kmeans25_distcount/kmeans25_count) %>%
  mutate(k50_prop = kmeans50_distcount/kmeans50_count) %>%
  mutate(k100_prop = kmeans100_distcount/kmeans100_count) %>%
  mutate(k200_prop = kmeans200_distcount/kmeans200_count) %>%
  select(-c('kmeans25_count','kmeans50_count','kmeans100_count','kmeans200_count','CountNorm')) %>%
  filter(k25_prop != 'inf')
```

Clean the imported predictor CSVs
```{r}
BioPhys_mean_clean = BioPhys_mean %>%
  dplyr::select(-c(.geo, PlotKey, SpeciesLis, VisitDate)) %>%
  dplyr::select(-1) 

clean_df <- function(df){
  df %>%
  dplyr::select(-c(PlotID, .geo, PlotKey, SpeciesLis, VisitDate, VisitYear,1))}

BioPhys_mode_clean <- clean_df(BioPhys_mode) %>% mutate(landform = mode) %>% dplyr::select(-mode)
Climate_ann_clean <- clean_df(Climate_ann)
Climate_mean_clean <- clean_df(Climate_mean)
Soil_clean <- clean_df(Soil)
RemSens_clust_clean <- clean_df(RemSens_clust) 
Fire_clean <- clean_df(Fire) %>% dplyr::select(-c(YearsOfMTBS,LastFire))%>%replace(., is.na(.), 0)

remove(BioPhys_mean,BioPhys_mode,Climate_ann,Climate_mean,Soil,Fire,RemSens,RemSens_clust)
```

Join the dataframes together in multiple ways
ALL PREDICTORS
```{r}
Predictors_all <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  left_join(Climate_mean_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  rename('PrimaryKey' = 'p.PrimaryKey', 'PlotID' = 'p.PlotID', 'VisitYear' = 'p.VisitYear') # revert names of non-predictors

#not sure where the 70 nas are coming from

#remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean)
```

REMOVING PREDICTORS FOR INDIVIDUAL BANDS
```{r}
Predictors_noBands <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean%>%select(-starts_with('B')), by = 'PrimaryKey') %>% #remove columns that start with 'B'
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  left_join(Climate_mean_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  rename('PrimaryKey' = 'p.PrimaryKey', 'PlotID' = 'p.PlotID', 'VisitYear' = 'p.VisitYear') # revert names of non-predictors
  
  #not sure where the 70 nas are coming from

#remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean)
```

REMOVING 30-YEAR MEAN CLIMATE DATA
```{r}
Predictors_noCliMn <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean, by = 'PrimaryKey') %>% #remove columns that start with 'B'
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  rename('PrimaryKey' = 'p.PrimaryKey', 'PlotID' = 'p.PlotID', 'VisitYear' = 'p.VisitYear') # revert names of non-predictors

#not sure where the 70 nas are coming from

#remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean)
```

REMOVING BANDS PREDICTORS AND 30-YEAR MEAN CLIMATE DATA
```{r}
Predictors_noCliMn_noBands <- BioPhys_mean_clean %>%
  left_join(BioPhys_mode_clean, by = 'PrimaryKey') %>%
  left_join(RemSens_clean%>%select(-starts_with('B')), by = 'PrimaryKey') %>% #remove columns that start with 'B'
  left_join(RemSens_clust_clean, by = 'PrimaryKey') %>%
  left_join(Soil_clean, by = 'PrimaryKey') %>%
  left_join(Fire_clean, by = 'PrimaryKey') %>%
  left_join(Climate_ann_clean, by = 'PrimaryKey') %>%
  drop_na() %>%
  setNames(paste0('p.', names(.))) %>% #add p prefix to denote predictors
  rename('PrimaryKey' = 'p.PrimaryKey', 'PlotID' = 'p.PlotID', 'VisitYear' = 'p.VisitYear') # revert names of non-predictors

#not sure where the 70 nas are coming from

remove(Fire_clean,Soil_clean,RemSens_clean,RemSens_clust_clean,BioPhys_mean_clean,BioPhys_mode_clean,Climate_mean_clean,Climate_ann_clean,clean_df)
```

# Import spatial plots and join with GEE zonal stats--largely unnecessary
```{r}
# library(sf)
# # Import plots sf for joining GEE zonal stats too
# AIM_plots <- read_sf('data/Shapefiles/AIM_model.shp') %>%
#   st_transform(crs = 4326)
# # Convert to tibble without geometry for manipulating and joining
# AIMdf <- as_data_frame(matrix(unlist(AIM_plots), nrow=nrow(AIM_plots), byrow=F))
# AIM_plotsDF <- AIMdf %>%
#   dplyr::select(-V3, -V7, -V8) %>%
#   dplyr::rename( "plotkey" = V1, "PlotID" = V2, "VisitYear" = V4, 
#                  "PrimaryKey" = V5, "SpeciesLis" = V6)
# 
# listofGEE = list(AIM_plotsDF,ndvi_mean,ndvi_std,nbr_mean,nbr_std,tcb_mean,tcb_std,tcg_mean,tcg_std,tcw_mean,tcw_std)
# # Join the zonal statistics to each plot, remove plots with NAs--likely only the dateless plot
# AIM_wGEE <- reduce(listofGEE, left_join)
```

### Import Species Diversity data
```{r}
Responses = read_csv('data/CSVs/ResponseVariables.csv') %>% setNames(paste0('r.', names(.))) %>% rename('PrimaryKey' = 'r.PrimaryKey')
```

Join the predictors and responses together
```{r}
BuildModelTable = function(preds){
  table <- preds %>%
            dplyr::select(-c(PlotID, VisitYear)) %>%
            full_join(Responses, 'PrimaryKey') %>%
            drop_na() 
  return(table)}

# All responses
ModelTable_all = BuildModelTable(Predictors_all)
# All predictors except individual bands
ModelTable_noBands = BuildModelTable(Predictors_noBands)
# All responses except 30-year climate means
ModelTable_noCliMn = BuildModelTable(Predictors_noCliMn)
ModelTable_noCliMn = subset(ModelTable_noCliMn, PrimaryKey %in% ModelTable_all$PrimaryKey)
# All responses except individual bands and 30-year climate means
ModelTable_noCliMn_noBands = BuildModelTable(Predictors_noCliMn_noBands)
ModelTable_noCliMn_noBands= subset(ModelTable_noCliMn_noBands, PrimaryKey %in% ModelTable_all$PrimaryKey)

remove(Predictors_all,Predictors_noBands,Predictors_noCliMn,Predictors_noCliMn_noBands)
```

################## Random Forest Modeling ######################
### Prepare Data for Model Selection
```{r}
### Generate random number to set seed
# sample(1:1000000,1)
rf_seed = 517953
set.seed(rf_seed)

## Subset dataframe into predictors and responses
subset_preds = function(ModelTable){
       predictors = ModelTable %>%
                       dplyr::select(starts_with('p.')) %>%
                       mutate(p.landform_v = factor(ModelTable$p.landform)) %>%
                       dplyr::select(-p.landform)
       return(predictors)}

Preds_All = subset_preds(ModelTable_all)
Preds_NoBands = subset_preds(ModelTable_noBands)
Preds_NoCliMn = subset_preds(ModelTable_noCliMn)
Preds_NoCliMn_NoBands = subset_preds(ModelTable_noCliMn_noBands)

subset_resp = function(ModelTable){
       responses = ModelTable %>%
                       dplyr::select(starts_with('r.'))
       return(responses)}

Resps_All = subset_resp(ModelTable_all)
remove(ModelTable_all,ModelTable_noBands,ModelTable_noCliMn,ModelTable_noCliMn_noBands,Responses)
```

Assess individual predictors against responses
```{r}
# Shannon_cor = t(data.frame(cor(responses$Shannon, predictors%>%dplyr::select(-landform_v))))
# InvSimp_cor = t(data.frame(cor(responses$InvSimp, predictors%>%dplyr::select(-landform_v))))
# Simpson_cor = t(data.frame(cor(responses$Simpson, predictors%>%dplyr::select(-landform_v))))
# Pielou_cor = t(data.frame(cor(responses$Pielou, predictors%>%dplyr::select(-landform_v))))
# renyi0_cor = t(data.frame(cor(responses$renyi0, predictors%>%dplyr::select(-landform_v))))
# renyi.5_cor = t(data.frame(cor(responses$renyi.5, predictors%>%dplyr::select(-landform_v))))
# renyi1_cor = t(data.frame(cor(responses$renyi1, predictors%>%dplyr::select(-landform_v))))
# renyi2_cor = t(data.frame(cor(responses$renyi2, predictors%>%dplyr::select(-landform_v))))
# renyi4_cor = t(data.frame(cor(responses$renyi4, predictors%>%dplyr::select(-landform_v))))
# renyi16_cor = t(data.frame(cor(responses$renyi16, predictors%>%dplyr::select(-landform_v))))
# renyi64_cor = t(data.frame(cor(responses$renyi64, predictors%>%dplyr::select(-landform_v))))
# hill0_cor = t(data.frame(cor(responses$hill0, predictors%>%dplyr::select(-landform_v))))
# hill.5_cor = t(data.frame(cor(responses$hill.5, predictors%>%dplyr::select(-landform_v))))
# hill1_cor = t(data.frame(cor(responses$hill1, predictors%>%dplyr::select(-landform_v))))
# hill2_cor = t(data.frame(cor(responses$hill2, predictors%>%dplyr::select(-landform_v))))
# hill4_cor = t(data.frame(cor(responses$hill4, predictors%>%dplyr::select(-landform_v))))
# hill16_cor = t(data.frame(cor(responses$hill16, predictors%>%dplyr::select(-landform_v))))
# hill64_cor = t(data.frame(cor(responses$hill64, predictors%>%dplyr::select(-landform_v))))
# SpecRich_cor = t(data.frame(cor(responses$SpecRich, predictors%>%dplyr::select(-landform_v))))
# PredNames = colnames(predictors[2:94])
# 
# options(scipen = 999)
# Cor_df = bind_cols(Preds=PredNames,Shan=Shannon_cor,Simp=Simpson_cor,ISimp=InvSimp_cor,Pielou=Pielou_cor,Ren0=renyi0_cor,Ren.5=renyi.5_cor,Ren1=renyi1_cor,Ren2=renyi2_cor,Ren4=renyi4_cor,Ren16=renyi16_cor,Ren64=renyi64_cor,hill0=hill0_cor,hill.5=hill.5_cor,hill1=hill1_cor,hill2=hill2_cor,hill4=hill4_cor,hill16=hill16_cor,hill64=hill64_cor, SpRi = SpecRich_cor)
# 
# corrplot(Cor_df)
# 
# remove(Shannon_cor,Simpson_cor,InvSimp_cor,Pielou_cor,renyi.5_cor,renyi0_cor,renyi16_cor,renyi1_cor,renyi2_cor,renyi4_cor,renyi64_cor,hill.5_cor,hill0_cor,hill1_cor,hill16_cor,hill2_cor,hill4_cor,hill64_cor,SpecRich_cor)
```

```{r}
# # Get mean correlation across all responses
# # Convert df to absolute values
# Corr_df_abs <- abs(Cor_df[2:20])
# predNames <- data.frame(Cor_df[1])
# predCor <- data.frame(rowMeans(Corr_df_abs))
# predRank <- cbind(predNames,predCor)
# View(predRank)
# 
# # Get mean correlation across all predictors
# colMeans(Corr_df_abs)
```

Random forest variable selection
```{r}
# rfselShan = rf.modelSel(xdata = predictors, ydata = responses$Shannon, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselSimp = rf.modelSel(xdata = predictors, ydata = responses$Simpson, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselISimp = rf.modelSel(xdata = predictors, ydata = responses$InvSimp, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselPielou = rf.modelSel(xdata = predictors, ydata = responses$Pielou, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen0 = rf.modelSel(xdata = predictors, ydata = responses$renyi0, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen.5 = rf.modelSel(xdata = predictors, ydata = responses$renyi.5, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen1 = rf.modelSel(xdata = predictors, ydata = responses$renyi1, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen2 = rf.modelSel(xdata = predictors, ydata = responses$renyi2, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen4 = rf.modelSel(xdata = predictors, ydata = responses$renyi4, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen16 = rf.modelSel(xdata = predictors, ydata = responses$renyi16, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselRen64 = rf.modelSel(xdata = predictors, ydata = responses$renyi64, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill0 = rf.modelSel(xdata = predictors, ydata = responses$hill0, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill.5 = rf.modelSel(xdata = predictors, ydata = responses$hill.5, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill1 = rf.modelSel(xdata = predictors, ydata = responses$hill1, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill2 = rf.modelSel(xdata = predictors, ydata = responses$hill2, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill4 = rf.modelSel(xdata = predictors, ydata = responses$hill4, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill16 = rf.modelSel(xdata = predictors, ydata = responses$hill16, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
# rfselHill64 = rf.modelSel(xdata = predictors, ydata = responses$hill64, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 

RFsel_All_SR = rf.modelSel(xdata = Preds_All, ydata = Resps_All$r.SpecRich, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
RFsel_NoBands_SR = rf.modelSel(xdata = Preds_NoBands, ydata = Resps_All$r.SpecRich, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
RFsel_NoCliMn_SR = rf.modelSel(xdata = Preds_NoCliMn, ydata = Resps_All$r.SpecRich, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
RFsel_NoCliMn_NoBands_SR = rf.modelSel(xdata = Preds_NoCliMn_NoBands, ydata = Resps_All$r.SpecRich, r = c(0.1, 0.2, 0.5, 0.7, 0.9)) 
```

Assess outputs from species richness model
```{r}
# plot(RFsel_All_SR)
# plot(RFsel_NoBands_SR)
# plot(RFsel_NoCliMn_SR)
# plot(RFsel_NoCliMn_NoBands_SR)

# Function to process the rf.modelSel() objects
Process_ModelSel = function(rfsel){
    rfsel <- cbind(rownames(rfsel$sel.importance), rfsel$sel.importance)
    rownames(rfsel) <- NULL
    colnames(rfsel) <- c("name", "imp")
    return(rfsel)}

RFsel_ALL_SR_ready = Process_ModelSel(RFsel_All_SR)
RFsel_NoBands_SR_ready = Process_ModelSel(RFsel_NoBands_SR)
RFsel_NoCliMn_SR_ready = Process_ModelSel(RFsel_NoCliMn_SR)
RFsel_NoCliMn_NoBands_SR_ready = Process_ModelSel(RFsel_NoCliMn_NoBands_SR)

###########################################################
######## Convert these to more succinct functions #########

# Order the dataframe by variable importance
RFsel_ALL_SR_ready <- RFsel_ALL_SR_ready[order(-RFsel_ALL_SR_ready$imp),]
RFsel_NoBands_SR_ready <- RFsel_NoBands_SR_ready[order(-RFsel_NoBands_SR_ready$imp),]
RFsel_NoCliMn_SR_ready <- RFsel_NoCliMn_SR_ready[order(-RFsel_NoCliMn_SR_ready$imp),]
RFsel_NoCliMn_NoBands_SR_ready <- RFsel_NoCliMn_NoBands_SR_ready[order(-RFsel_NoCliMn_NoBands_SR_ready$imp),]

# Get list of names of top 30 predictors for each model
RFsel_ALL_SR_names <- RFsel_ALL_SR_ready$name[1:nrow(RFsel_ALL_SR_ready)]
RFsel_NoBands_SR_names <- RFsel_NoBands_SR_ready$name[1:nrow(RFsel_NoBands_SR_ready)]
RFsel_NoCliMn_SR_names <- RFsel_NoCliMn_SR_ready$name[1:nrow(RFsel_NoCliMn_SR_ready)]
RFsel_NoCliMn_NoBands_SR_names <- RFsel_NoCliMn_NoBands_SR_ready$name[1:nrow(RFsel_NoCliMn_NoBands_SR_ready)]

# drop columns that are not as important based on rfmodelsel--columns not in list of names
RFsel_ALL_SR_cov <- Preds_All[,as.character(RFsel_ALL_SR_names)] #[-c(4,5,6,9,8,7)]]
RFsel_NoBands_SR_cov <- Preds_NoBands[,as.character(RFsel_NoBands_SR_names)]
RFsel_NoCliMn_SR_cov <- Preds_NoCliMn[,as.character(RFsel_NoCliMn_SR_names)]
RFsel_NoCliMn_NoBands_SR_cov <- Preds_NoCliMn_NoBands[,as.character(RFsel_NoCliMn_NoBands_SR_names)]

#Too many GD object--get rid of some
remove(RFsel_ALL_SR_ready,RFsel_NoBands_SR_ready,RFsel_NoCliMn_NoBands_SR_ready,RFsel_NoCliMn_SR_ready,Predictors_all,Predictors_noBands,Predictors_noCliMn,Predictors_noCliMn_noBands,BuildModelTable,Process_ModelSel,subset_preds,subset_resp)
```

Assess correlation between model predictors
```{r}
################################################################################
########### Figure out better way to eliminate correlated variables ############

# calculate correlation coefficient matrix
# Figure out how to include landform
cor_All <-cor(RFsel_ALL_SR_cov%>%select(-p.landform_v), method="pearson")
cor_NoBands <-cor(RFsel_NoBands_SR_cov, method="pearson")
cor_NoCliMn <-cor(RFsel_NoCliMn_SR_cov%>%select(-p.landform_v), method="pearson")
cor_NoCliMn_NoBands <-cor(RFsel_NoCliMn_NoBands_SR_cov%>%select(-p.landform_v), method="pearson")

# plot the correlation. the darker the number, the more correlated the two
# variables
corrplot(cor_All,method="number")
corrplot(cor_NoBands,method="number")
corrplot(cor_NoCliMn,method="number")
corrplot(cor_NoCliMn_NoBands,method="number")

rastercolumns_full <- predictors[,as.character(raster_cov_names)]
rastercolumns_6var <- predictors[,as.character(raster_cov_names)[-c(7:14)]]
rastercolumns_7var <- predictors[,as.character(raster_cov_names)[-c(7,8,9,11,12,13,14)]]

rfselFull$selvars
## You can now manually remove any variables that are above your correlation 
## threshold. In this case we will just leave them all in, but we could remove
## them by using the following code

###### Model fitting
ALL_SR_model <- cbind(RFsel_ALL_SR_cov, specN= Resps_All$r.SpecRich)
NoBands_SR_model <- cbind(RFsel_NoBands_SR_cov, specN= Resps_All$r.SpecRich)
NoCliMn_SR_model <- cbind(RFsel_NoCliMn_SR_cov, specN= Resps_All$r.SpecRich)
NoCliMn_NoBands_model <- cbind(RFsel_NoCliMn_NoBands_SR_cov, specN= Resps_All$r.SpecRich)

# Now let's restrict our final predictors to those that are most important (Change manually) 
# and run the continuous model
```

Run random forest models
```{r}
ALL_rf_model = randomForest(specN ~ ., data=ALL_SR_model, importance = TRUE, ntree = 3000, mtry = 2)
NoBands_rf_model = randomForest(specN ~ ., data=NoBands_SR_model, importance = TRUE, ntree = 3000, mtry = 2)
NoCliMn_rf_model = randomForest(specN ~ ., data=NoCliMn_SR_model, importance = TRUE, ntree = 3000, mtry = 2)
NoCliMn_NoBands_rf_model = randomForest(specN ~ ., data=NoCliMn_NoBands_model, importance = TRUE, ntree = 3000, mtry = 2)

# Use the strongest model going forward
NoBands_rf_model # 46.05% variance explained!!!

############# Evaluate this code ###################
####################################################
NoBands_rf_model$rsq #Pseudo r-squared of .459
plot(NoBands_rf_model$rsq)
varImpPlot(NoBands_rf_model[[1]])
imp_test[[1]]
```

Plot the predicted data vs. the observed data
```{r}
# Round for later plotting
NoBands_rf_modelEx <- round(NoBands_rf_model$rsq[3000], digits = 4)

NoBands_predicted <- NoBands_rf_model$predicted

# plot observed values on the x-axis and predicted values on the y-axis. we are
# looking for these to be correlated (so close to a 1:1 line)
# Plot as Alpha ggplot
NoBands_assess <- bind_cols(predict = NoBands_predicted, response = Resps_All$r.SpecRich)

varNoBands <- text_grob(paste("Variance explained =", as.character(NoBands_rf_modelEx[1:4]), sep = "\n"), x = .2, y = .9)

gg_NoBands <- ggplot(NoBands_assess, aes(x=response, y = predict)) +
  geom_point(alpha=.2, color = 'blue') +
  geom_smooth(method='lm',formula=y~x, color = 'red') +
  labs(x = "Observed diversity", y = "Predicted diversity",title="28 variables, 1000 trees") +
  annotation_custom(varNoBands) +
  theme_few()
gg_NoBands

# calculate r-squared and RMSE
actual = NoBands_assess$response
predicted = NoBands_assess$predict
R2 <- 1 - (sum((actual-predicted)^2)/sum((actual-mean(actual))^2))
library(caret)
rmse <- RMSE(NoBands_assess$predict,NoBands_assess$response)

# model_comp <- egg::ggarrange(gg1000var,gg3000var,gg5000var, nrow = 1)
# annotate_figure(model_comp, top = text_grob("Comparing models with 6 and 14 variables", face = 'bold', size = 18))
```

Evaluate Error
```{r}
# Calculate RMSE for the models
error6 <- predicted6 - response
error7 <- predicted7 - response
errorFull <- predictedFull - response

sqError6 <- error6^2
sqError7 <- error7^2
sqErrorFull <- errorFull^2

rmse <- function(error)
{
  sqrt(mean(error^2))
}

RMSE6 <- rmse(error6)
RMSE7 <- rmse(error7)
RMSEfull <- rmse(errorFull)

AIM_error6 <- bind_cols(error6 = error6,sqError6 = sqError6)
AIM_error7 <- bind_cols(error7 = error7,sqError7 = sqError7)
AIM_errorFull <- bind_cols(errorFull = errorFull,sqErrorFull = sqErrorFull)
AIM_error <- AIM_model %>%
  dplyr::select(1:2,32) %>%
  bind_cols(AIM_error6) %>%
  bind_cols(AIM_error7) %>%
  bind_cols(AIM_errorFull)

class(AIM_error)
View(AIM_error)
AIM_error_spatial <- st_sf(AIM_error)
HighError <- AIM_error_spatial %>%
  filter(errorFull >= 12 | errorFull <= -12) 
# 690 entries
RandomSample <- AIM_error[sample(1:nrow(AIM_error), nrow(HighError),
   replace=FALSE),]
RandomSampleSpat <-st_sf(RandomSample)

# Plot KDEs for each
highSP <- as(HighError, "Spatial")
highKDE <- sp.kde(testSP,n = 5000, bw = .4)
plot(highKDE)

writeRaster(highKDE,filename = 'data/Tiffs/highErrorKDE.tif', format = "GTiff")
st_write(HighError, 'data/Working/HighError.shp')

randomSP <- as(RandomSampleSpat, "Spatial")
randomKDE <- sp.kde(randomSP,n = 5000, bw = .4)
plot(randomKDE)

writeRaster(randomKDE,  filename = 'data/Tiffs/randomSampleKDE.tif', format = "GTiff")
st_write(RandomSample, 'data/Working/RandomSample.shp')

mapview(HighError,cex = 3, zcol = 'errorFull')

mapview(RandomSampleSpat,cex = 3, zcol = 'errorFull')
```



### Load GEE 2017 data for predictive mapping
```{r}
ndvi2017<- raster('data/Tiffs/GEE2017/ndvi2017.tif')
nbr2017 <- raster('data/Tiffs/GEE2017/nbr2017.tif')
tcb2017 <- raster('data/Tiffs/GEE2017/tcb2017.tif')
tcg2017 <- raster('data/Tiffs/GEE2017/tcg2017.tif')
tcw2017 <- raster('data/Tiffs/GEE2017/tcw2017.tif')
plot(ndvi2017)
plot(nbr2017)
plot(tcb2017)
plot(tcg2017)
plot(tcw2017)

# Import independent validation data
AIM_covariates <- read_sf('data/PlotsShapefiles/AIM_validate.shp') %>%
  st_transform(crs = 4326)
View(AIM_covariates)


testPredict <- predict(rf_fit1000)
### Reports accuracy, Cohen's kappa and area under ROC curve
testAssessPred <- accuracy(x = rf_fit1000, resp)
```

```{r}
# Import shapefile to clip
GBsubset <- read_sf('data/Shapefiles/ProjectSubset.shp') %>%
  st_transform(crs = 4326)

# Rasters needed to predict
tcb <- raster('data/Tiffs/GEE2017/tcb2017.tif')
tcg <- raster('data/Tiffs/GEE2017/tcg2017.tif')
cli_ffp <- raster('data/Tiffs/Cli/climate_ffp.tif')
trasp <- raster('data/Tiffs/Phys/GBtrasp.tif')
soil_h20 <- raster('data/Tiffs/Soil/SoilWaterMosaic.tif')

# Clip rasters to GB bounds
tcb_clip <- crop(tcb, GBsubset)
tcg_clip <- crop(tcg, GBsubset)
cli_ffp_clip <- crop(cli_ffp, GBsubset)
trasp_clip <- crop(trasp, GBsubset)
soil_h20_clip <- crop(soil_h20, GBsubset)
plot(tcb_clip)
plot(tcg_clip)
plot(cli_ffp_clip)
plot(trasp_clip)
plot(soil_h20_clip)

trasp_clip$extent
tcb_clip
tcg_clip
cli_ffp_clip
soil_h20_clip

mapview(GBsubset)

extent_proj = st_transform(extent,crs = 4326)
mapview(extent)

tcb_proj <- projectRaster(tcb_clip, crs="+proj=longlat +datum=WGS84")
tcg_proj <- projectRaster(tcg_clip, crs="+proj=longlat +datum=WGS84")
cli_ffp_proj <- projectRaster(cli_ffp_clip, crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
trasp_proj <- projectRaster(trasp_clip, crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
soil_h20_proj <- projectRaster(soil_h20_clip, crs="+proj=longlat +datum=WGS84")

stack_5band=stack(tcb_clip, tcg_clip, cli_ffp_clip, clip, clip)

latitude

```
