---
title: "SpeciesDiversityIndices"
author: "Eric Jensen"
date: "November 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse) 
library(reshape)
library(reshape2)
library(sf)
library(BiodiversityR)
library(rlang)
```

# Read in LPI CSVs as tibbles
```{r}
options(scipen=999)

# Read in AIM database data
LPI_detail <- read_csv('data/CSVs/Full_LPIDetailTable.csv') %>%
  dplyr::select(-OBJECTID)

# Read in 2019 summer field data
Jensen_LPI_header <- read_csv('data/CSVs/Jensen_tblLPIHeader.csv')
#Jensen_LPI_detail <- read_csv('data/CSVs/Jensen_tblLPIDetail.csv') #-- This dataframe is needed for the below code chunk
Jensen_LPI_detail <- read_csv('data/CSVs/Jensen_tblLPIDetail_wPlotKey.csv')

# Spatially subset plots to Great Basin based on keys in LPI_detail that spatially subset to great basin
GBbounds <- read_sf('data/Shapefiles/ProjectBounds.shp') %>%
  st_transform(crs = 2163)
AIMfull <- read_sf('data/Shapefiles/AIM_tblPlots.shp') %>%
  st_transform(crs = 2163)

AIMfull_keys = dplyr::select(AIMfull, PrimaryKey)

AIMgb_keys <- AIMfull_keys[GBbounds,] #Spatially subset keys to Great Basin

LPI_detail_GB = subset(LPI_detail, PrimaryKey %in% AIMgb_keys$PrimaryKey) #Keep only rows that match Great Basin keys
# LPI_detail_GB will be the dataframe that I will use going forward
remove(GBbounds,AIMfull,LPI_detail, AIMgb_keys, AIMfull_keys)
```

### Need to get PlotKeys for the Jensen_LPI_detail data to assign as primary keys (as done previously in WithholdMergeAIM script)
```{r}
# # Isolate rec and line keys and join to the Jensen_LPI_detail data
# RecLineKeys <- Jensen_LPI_header %>%
#   select(RecKey,LineKey)
# joinedDetail_rec <- left_join(Jensen_LPI_detail, RecLineKeys, by = "RecKey")
# 
# # Import tbl Lines, Isolate the line and plot keys and join to the Jensen_LPI_detail data
# Jensen_lines <- read_csv('data/CSVs/Jensen_tblLines.csv')
# LinePlotKeys <- Jensen_lines %>%
#   select(LineKey,PlotKey)
# joinedDetail_recPlot <- left_join(joinedDetail_rec, LinePlotKeys, by = "LineKey")
# 
# # Export LPI_detail table as Jensen_TblLPIDetail_wPlotKey.csv
# renamed_clean <- joinedDetail_recPlot %>%
#   rename(SoilSurfac = SoilSurface, ChkboxLowe  = ChkboxLower1, ChkboxLo_1 = ChkboxLower2, ChkboxLo_2 = ChkboxLower3, ChkboxLo_3 = ChkboxLower4, HeightLowe = HeightLower1, HeightLo_1 = HeightLower2, HeightLo_2 = HeightLower3, HeightLo_3 = HeightLower4, HeightSurf = HeightSurface, HeightWood = HeightWoody, HeightHerb = HeightHerbaceous, SpeciesWoo = SpeciesWoody, SpeciesHer = SpeciesHerbaceous, ChkboxWood = ChkboxWoody, ChiboxHerb = ChkboxHerbaceous, ChkboxLo_4 = ChkboxLower5, ChkboxLo_5 = ChkboxLower6, ChkboxLo_6 = ChkboxLower7, HeightLo_4 = HeightLower5, HeightLo_5 = HeightLower6, HeightLo_6 = HeightLower7, PrimaryKey = PlotKey)
# View(renamed_clean)
# 
# write_csv(renamed_clean, 'data/CSVs/Jensen_TblLPIDetail_wPlotKey.csv')
# remove(renamed_clean,joinedDetail_rec,joinedDetail_recPlot,Jensen_lines,RecLineKeys,LinePlotKeys)
# # Done -- This should remain commented out as a record of my workflow
```


### Clean the LPI data so that they can be combined
```{r}
# Convert logical columns to 0 and 1 integers
Jensen_LPI_detail <- Jensen_LPI_detail %>%
  mutate(ChkboxTop = as.numeric(ChkboxTop))%>%
  mutate(ChkboxLowe = as.numeric(ChkboxLowe))%>%
  mutate(ChkboxLo_1  = as.numeric(ChkboxLo_1))%>%
  mutate(ChkboxLo_2 = as.numeric(ChkboxLo_2))%>%
  mutate(ChkboxLo_3 = as.numeric(ChkboxLo_3))%>%
  mutate(ChkboxSoil = as.numeric(ChkboxSoil))%>%
  mutate(ChkboxLo_4 = as.numeric(ChkboxLo_4))%>%
  mutate(ChkboxLo_5 = as.numeric(ChkboxLo_5))%>%
  mutate(ChkboxLo_6 = as.numeric(ChkboxLo_6))%>%
  mutate(ChkboxWood = as.numeric(ChkboxWood))%>%
  mutate(ChiboxHerb = as.numeric(ChiboxHerb))%>%
  mutate(ChkboxLowerHerb = as.numeric(ChkboxLowerHerb))%>%
  mutate(PrimaryKey = as.character(PrimaryKey)) %>%
  mutate(Source = "FieldSeason2019") # Data source column

LPI_detail_GB <- mutate(LPI_detail_GB, Source = "AIMDataset") # Data source column
# View(LPI_detail)
# Bind rows to produce dataframe of all observations
# Soil surface hits are not included because if a species is a basal hit then it will also be included in the point as an upper point hit.
LPI_combined <- Jensen_LPI_detail %>%
  bind_rows(LPI_detail_GB) %>% 
  dplyr::select(PointNbr,TopCanopy,Lower1,Lower2,Lower3,Lower4,Lower5,Lower6,Lower7,LineKey,PrimaryKey,Source)
#View(LPI_combined)
```

### Convert the dataframe from wide to long

#### Filter out all "None" points
```{r}
# Melt the dataframe from wide to long
MeltLPI = melt(LPI_combined, id = c('PointNbr','LineKey','PrimaryKey','Source'))
#View(MeltLPI)

# Assess the outcome
noNone_strings = MeltLPI %>%
  dplyr::select(value) %>%
  filter(value != 'None') %>%
  group_by(value) %>%
  mutate('count' = n()) %>%
  distinct()
#View(noNone_strings)

MeltLPI_woNone = filter(MeltLPI, value != 'None') 
#View(MeltLPI_woNone)
remove(MeltLPI,noNone_strings, LPI_combined)
```

#### Assess codes of 3 characters or fewer
```{r}
# View distinct values for species
# Joining the USDA species codes will help with QC
# Filter by number of characters and the count for each
# Need to remove non-species hits somehow
short_strings = MeltLPI_woNone %>%
  dplyr::select(value) %>%
  filter(!nchar(value) >3) %>%
  group_by(value) %>%
  mutate('count' = n()) %>%
  distinct()
#View(short_strings)

#Based on this assessment I will remove all observations of three characters or fewer--See "Processing AIM Species Data" Google Doc for justification, removed below

MeltLPI_woShort = filter(MeltLPI_woNone, nchar(value) >3) 
#View(MeltLPI_woShort)
remove(MeltLPI_woNone,short_strings)
```

#### Remove all unknown plant codes
```{r}
# Isolate and count unknown plants codes in the LPI dataset
unk_codes = c('^AF0','^AF1','^AF2','^AF3','^AF4','^AF5','^AF6','^AF7','^AF8','^AF9',
              '^PF0','^PF1','^PF2','^PF3','^PF4','^PF5','^PF6','^PF7','^PF8','^PF9',
              '^AG0','^AG1','^AG2','^AG3','^AG4','^AG5','^AG6','^AG7','^AG8','^AG9', 
              '^PG0','^PG1','^PG2','^PG3','^PG4','^PG5','^PG6','^PG7','^PG8','^PG9',
              '^SH0','^SH1','^SH2','^SH3','^SH4','^SH5','^SH6','^SH7','^SH8','^SH9',
              '^AAFF','^AFAF','^AAGG','^AGAG','^PPFF','^PFPF','^PPGG','^PGPG','^SHSH', '^PPSH', '^PPGG')

unk_strings = MeltLPI_woShort %>%
  dplyr::select(value) %>%
  filter(grepl(paste(unk_codes,collapse='|'),value)) %>%
  group_by(value) %>%
  mutate('count' = n()) %>%
  distinct()
#View(unk_strings)

# Compare number of unknowns with total observations in the datebase
## Sum the total number of points in the database
total_obs = MeltLPI_woShort %>%
  dplyr::select(value) %>%
  group_by(value) %>%
  mutate('count' = n()) %>%
  distinct()
#View(total_obs)

sum_total = sum(total_obs$count)
sum_unk = sum(unk_strings$count)
sum_unk / sum_total
## Unknowns comprise just over 1.2 percent of all points

MeltLPI_ready = filter(MeltLPI_woShort, !grepl(paste(unk_codes,collapse='|'),value)) 
#View(MeltLPI_ready)

remove(total_obs,unk_strings,unk_codes,MeltLPI_woShort,sum_unk, sum_total,Jensen_LPI_detail,Jensen_LPI_header)
```
#### Note: Data are now cleaned enough to work directly with the species data as observations and to join and compare them with the PLANTS database

### Assess remaining observations in the context of the USDA PLANTS database--CAN SKIP THIS FOR PRODUCING INDICES!!
#### Join LPI codes to USDA PLANTS codes
```{r}
# Read in USDA plants dataframe to get genus and species data
# I combined the complete USDA PLANTS list with the Nevada, Oregon, Idaho, Utah, and California PLANTS list to create the below list
plantsList <- read_csv('data/TXT/Combined_PlantsTable_wAdditions.txt') %>%
  dplyr::rename('SynSym' = 'Synonym Symbol') %>%
  dplyr::rename('LatinName' = 'Scientific Name with Author') %>%
  dplyr::rename('CommonName' = 'Common Name') %>%
  distinct()
#View(plantsList)

# Remove rows of plants list with synonym symbols and keep only those with NA
plantsList = filter(plantsList,is.na(SynSym)) 
#View(plantsList)

# Join the USDA plants data to the value column
lpi_plants_join = MeltLPI_ready %>%
  select(value) %>%
  dplyr::rename(Symbol = value) %>%
  group_by(Symbol) %>%
  mutate('count' = n()) %>%
  left_join(plantsList, 'Symbol') 
#View(lpi_plants_join)

# Create dataframe of joined species
lpi_plants_joined = filter(lpi_plants_join, !is.na(LatinName))
#View(lpi_plants_joined)
```

#### Match NA codes to Synonym Symbols, as possible
```{r}
# Filter down to those rows that were not joined previously
lpi_plants_unjoined = lpi_plants_join %>%
  filter(is.na(LatinName))
#View(lpi_plants_unjoined)

# Reimport PLANTS database and keep the Species Synonyms
plantsList_wSyn <- read_csv('data/TXT/Combined_PlantsTable_wAdditions.txt') %>%
  dplyr::rename('SynSym' = 'Synonym Symbol') %>% 
  dplyr::rename('LatinName' = 'Scientific Name with Author') %>%
  dplyr::rename('CommonName' = 'Common Name') %>%
  distinct()
#View(plantsList_wSyn)

# Join the unjoined LPI observations to the Synonym Symbols
lpi_plants_unjoined_joined = lpi_plants_unjoined %>%
  select(Symbol, count) %>%
  left_join(plantsList_wSyn, c('Symbol' = 'SynSym')) %>%
  filter(!is.na(LatinName)) %>%
  dplyr::rename('SynSym' = 'Symbol.y') %>%
  distinct()
#View(lpi_plants_unjoined_joined)

# Bind joined rows from "SynSym" to the above rows joined to "Symbol"
lpi_plants_bind <- rbind(lpi_plants_joined, lpi_plants_unjoined_joined)
#View(lpi_plants_bind)
```

#### Assess remaining unjoined plants
```{r}
lpi_still_na = lpi_plants_unjoined %>%
  select(Symbol, count) %>%
  left_join(plantsList_wSyn, c('Symbol' = 'SynSym')) %>%
  filter(is.na(LatinName))
#View(lpi_still_na)
remove(lpi_plants_unjoined, lpi_plants_unjoined_joined)
# Have reduced the number of unknown species significantly! Need to figure out how to treat these though
```


#########################################################################################################
### Need to figure out data back-up system--be an adult. https://www.crashplan.com/en-us/pricing/ #######
#########################################################################################################

### Apply the USDA PLANTS joins to the full dataset of plants--RESUME CALCULATIONS HERE!!
```{r}
# Read in USDA plants dataframe to get genus and species data
# I combined the complete USDA PLANTS list with the Nevada, Oregon, Idaho, Utah, and California PLANTS list to create the below list
plantsList <- read_csv('data/TXT/Combined_PlantsTable_wAdditions.txt') %>%
  dplyr::rename('SynSym' = 'Synonym Symbol') %>%
  dplyr::rename('LatinName' = 'Scientific Name with Author') %>%
  dplyr::rename('CommonName' = 'Common Name') %>%
  distinct()
#View(plantsList)

# Remove rows of plants list with synonym symbols and keep only those with NA
plantsList = filter(plantsList,is.na(SynSym)) 
#View(plantsList)

# Join the USDA plants data to the value column
lpi_USDA_full_join = MeltLPI_ready %>%
  dplyr::rename(Symbol = value) %>%
  left_join(plantsList, 'Symbol') 
#View(lpi_USDA_full_join)

# Create dataframe of joined species
lpi_USDA_full_joined = filter(lpi_USDA_full_join, !is.na(LatinName))
#View(lpi_plants_joined)
```

#### Match NA codes to Synonym Symbols, as possible
```{r}
# Filter down to those rows that were not joined previously
lpi_USDA_full_unjoined = lpi_USDA_full_join %>%
  filter(is.na(LatinName)) %>%
  dplyr::select(-c(SynSym, LatinName,CommonName,Family))
#View(lpi_USDA_full_unjoined)

# Reimport PLANTS database and keep the Species Synonyms
plantsList_wSyn <- read_csv('data/TXT/Combined_PlantsTable_wAdditions.txt') %>%
  dplyr::rename('SynSym' = 'Synonym Symbol') %>% 
  dplyr::rename('LatinName' = 'Scientific Name with Author') %>%
  dplyr::rename('CommonName' = 'Common Name') %>%
  distinct()
#View(plantsList_wSyn)

# Join the unjoined LPI observations to the Synonym Symbols
lpi_USDA_full_unjoined_joined = lpi_USDA_full_unjoined %>%
  left_join(plantsList_wSyn, c('Symbol' = 'SynSym')) %>%
  filter(!is.na(LatinName)) %>%
  dplyr::rename('SynSym' = 'Symbol.y') %>%
  dplyr::select(-Family) %>%
  mutate("Family" = NA)  #remove family because of issues with duplication, can add later
#View(lpi_USDA_full_unjoined_joined)
#View(lpi_USDA_full_joined)

# Bind joined rows from "SynSym" to the above rows joined to "Symbol"
LPI_FULLY_JOINED <- rbind(lpi_USDA_full_joined, lpi_USDA_full_unjoined_joined)
#View(lpi_plants_bind)

# Count number of unjoined rows and as a percentage of the total points that we tried to be joined
#nrow(lpi_USDA_full_unjoined)-nrow(lpi_USDA_full_unjoined_joined)
# 3550, not bad!
#(nrow(lpi_USDA_full_unjoined)-nrow(lpi_USDA_full_unjoined_joined)) / ((nrow(lpi_USDA_full_unjoined)-nrow(lpi_USDA_full_unjoined_joined)) + nrow(lpi_plants_bind))
#Less than 3/10 of a percent
remove(lpi_USDA_full_join, lpi_USDA_full_joined, lpi_USDA_full_unjoined, lpi_USDA_full_unjoined_joined, plantsList, plantsList_wSyn)
```

---------------------------------------------------------------------------------
### Calculate Species Diversity Indices
```{r}
#####################################################
################ With test data #####################

# Apply the LPI_FULLY_JOINED dataframe here

# Notes:
# 1) Need to deal with species only ID'd to genus!!!
# 2) Need to incorporate Species Richness counts too!!!
# 3) Need to account for different plotsizes--rarefaction?--different sizes but same number of points; would affect species richness more than LPI/abundance

#BiodiversityR will be a useful package!!
#Vegan package as well

# data("BCI")
# #View(BCI) # Format AIM data in this way
# 
# # Assess rarefaction curves, chao indices, etc.
# 
# BCI_shan <- diversity(BCI, index = "shannon")
# BCI_simp <- diversity(BCI, index = "simpson")
# BCI_invSimp <- diversity(BCI, index = "invsimpson")
```

### Prepare data to calculate diversity 
```{r}
#View(LPI_FULLY_JOINED)

group_cols <- c('PrimaryKey', 'Symbol')


#View(Plots)

# Create dataframe of counts in a matrix form to calculate species diversity indices by
Species_Calcs <- LPI_FULLY_JOINED %>%
  dplyr::select(PrimaryKey,Symbol) %>%
  group_by(!!!syms(group_cols)) %>% # Group by both columns to get accuate counts of abundance for each plot
  mutate('Count' = n()) %>%
  distinct() %>%
  spread(Symbol, Count, fill = 0)

# Assess the dataframe contents
# Species_Calcs %>%
#   slice(1:100) %>%
#   select(1:20) %>%
#   View()
```

### Calculate diversity indices and join them to main dataframe
```{r}
# Calculate indices for each plot
shannon <- as_tibble(diversity(Species_Calcs[,2:length(Species_Calcs)], index = "shannon"))
simpson <- as_tibble(diversity(Species_Calcs[,2:length(Species_Calcs)], index = "simpson"))
invsimp <- as_tibble(diversity(Species_Calcs[,2:length(Species_Calcs)], index = "invsimpson"))

# Calculate Renyi and Hill Indices
# Supporting document for vegan package: https://cran.r-project.org/web/packages/vegan/vignettes/diversity-vegan.pdf
hill = renyi(Species_Calcs[,2:1340], scales = c(0,.5, 1, 2, 4, 16, 64), hill = TRUE) %>% 
  dplyr::rename(hill0=`0`,hill.5=`0.5`,hill1=`1`,hill2=`2`,hill4=`4`,hill16=`16`,hill64=`64`)
renyi = as_tibble(renyi(Species_Calcs[,2:1340], scales = c(0,.5, 1, 2, 4, 16, 64))) %>% 
  dplyr::rename(renyi0=`0`,renyi.5=`0.5`,renyi1=`1`,renyi2=`2`,renyi4=`4`,renyi16=`16`,renyi64=`64`)

# Pielou's evenness
S <- specnumber(Species_Calcs) ## rowSums(BCI > 0) does the same...
pielou <- shannon/log(S)

# Bind columns of indices to columns of plots
Plots <- LPI_FULLY_JOINED %>% # Produce dataframe with one column of plots to bind to
  dplyr::select(PrimaryKey) %>%
  distinct() 

PlotsIndices <- Plots %>%
  mutate('InvSimp' = invsimp$value) %>%
  mutate('Shannon' = shannon$value) %>%
  mutate('Simpson' = simpson$value) %>%
  mutate('Pielou' = pielou$value) %>%
  bind_cols(renyi) %>%
  bind_cols(hill) 

# Plots_wCounts <- LPI_FULLY_JOINED %>% # Produce dataframe with only Keys, Plant symbols, and counts of plants at each plot
#   dplyr::select(PrimaryKey,Symbol) %>%
#   group_by(!!!syms(group_cols)) %>% 
#   mutate('Count' = n()) %>%
#   distinct()
# 
# # Join the diversity indices to the plots with counts dataframe
# Plots_wCountsIndices <- Plots_wCounts %>%
#   inner_join(PlotsBound, by = 'PrimaryKey')

View(Plots_wCountsIndices)
```

Also consider:
Chao richness
Inverse Berger-Parker
Fisher's alpha
Brillouin
Margalef
Q-Statistic
Hill's evenness

Get species richness
```{r}
#Import shapefile with species richness counts
AIM_covariates <- sf::read_sf('data/Working/AIM_covariates_model.shp') %>%
  dplyr::select(c(PrimaryKey,SpecRich=7))

# Join to plots Indices
AllIndices <- PlotsIndices %>%
  left_join(AIM_covariates, 'PrimaryKey') %>%
  dplyr::select(-geometry)%>%
  drop_na()
```

Write species diversity indices to CSV
```{r}
write_csv(AllIndices, path = 'data/CSVs/ResponseVariables.csv')
```






